### 2017-11-26 notes_jc.tex
some notes

\documentclass[12pt]{article}
\usepackage{jchen2}

\begin{document}
\begin{center}
\begin{tabular}{cc}\toprule
    Notation & Explanation \\
    \midrule
    $w_i$ & Weight of $i$th grid node \\
    $\bm W$ & $\{w_1,\ldots, w_N\}$ \\
    $Z\uppr{k} = \{i_1\uppr{k}, \ldots, i_{n\uppr{k}}\uppr{k}\}$ & Path taken by trip $k$ \\
    $x_1\uppr{k}, x_2\uppr{k}$ & Start and end locations of trip $k$\\
    $\bm Z\uppr{k} = \bm Z(x_1\uppr{k},x_2\uppr{k})$ & Set of reasonable paths for trip $k$\\
    $T\uppr{k}$ & Time taken by trip $k$ \\\bottomrule
\end{tabular}
\end{center}

Assume the following model \begin{align*}
\bm W &\sim p_{\bm W} \\
Z\uppr{k} | \bm W &\sim p_{Z | \bm W} \tag{a distribution over $\bm Z\uppr{k}$}\\
T\uppr{k} | Z\uppr{k}, \bm W &\sim \Norm\pr{\sum_{i\in Z\uppr{k}} w_i, \sigma^2}
\end{align*}
We are interested in the posterior distribution of $\bm W$: 
\begin{align*}
p(\bm W | T\uppr{k}) &= \frac{p(T\uppr{k}| \bm W) p(\bm W)}{p(T\uppr{k})} \\
&=\frac{\bk{\sum_{Z\uppr{k} \in \bm Z\uppr{k}} p(T\uppr{k}|Z\uppr{k}, \bm W) p(Z\uppr{k}| \bm W)} p(\bm W)}{p(T\uppr{k})}
\end{align*}

If we are only interested in the MAP, then we have \[
\log p(\bm W | T\uppr{k}) = \underbrace{\log \pr{\sum_{Z\uppr{k} \in \bm Z\uppr{k}} p(T\uppr{k}|Z\uppr{k}, \bm W) p(Z\uppr{k}| \bm W)}}_{\ell(\bm W | T\uppr{k}) \text{ log-likelihood}} + \underbrace{\log p(\bm W)}_{\text{log prior}} + \text{const.}
\]

Our main challenge here is to compute the log-likelihood, which involves a sum over all of $\bm Z\uppr{k}$. If the set of reasonable paths include all paths inside a rectangle of size $n \times m$, then we have $\binom{n+m}{n} \sim \frac{(n+m)^n}{n^n}e^{n}$ terms, which is clearly intractable. Zhan, et al. solves the issue by computing the 20 shortest path as the reasonable path set and then running softmax.

\paragraph{Can we EM?} EM algorithm approximates the log-likelihood via 
\begin{align*}
Q(\bm W|\bm W^{(t)}) &= E\bk{\log p(T\uppr{k}, Z\uppr{k} | \bm W) | T\uppr{k}, \bm W\uppr{t}} \\&= -\frac{1}{2\sigma^2}\sum_{Z\uppr{k} \in \bm Z\uppr{k}} \pr{T\uppr{k} - \sum_{Z\uppr{k}} w_i}^2 p(Z\uppr{k} | T\uppr{k}, \bm W\uppr{t}) + \text{const.}
\end{align*}
However, (1) we again have to evaluate a sum of order $|\bm Z|$ and (2) we don't have good estimates of $p(Z\uppr{k} | T\uppr{k}, \bm W\uppr{t})$.

\paragraph{Can we Monte Carlo?} We might consider using rejection sampling to approximate \[
E_{Z}\bk{p(T\uppr{k}| Z\uppr{k}, \bm W)|\bm W} \approx \frac{1}{L}\sum_{j=1}^L p(T\uppr{k}| \tilde Z\uppr{k}_j, \bm W).
\]
The problem here is that the gradient update via \texttt{autograd} cannot take into account the gradient incurred \emph{in the sampling process}.

\paragraph{Can we variational inference?} Maybe. But (1) mean field doesn't really apply since $Z\uppr{k}$ cannot be broken into independent variables naturally and (2) VAE a la Kingma and Welling only works on continuous latent variables.

\paragraph{A Monte Carlo based method} We repeatedly sample $\tilde Z$ \emph{uniformly} from $\bm Z$ and use the sample mean to approximate summations over $\bm Z$. \begin{align*}
&\log E_Z\pr{p(T\uppr{k} | Z\uppr{k}, \bm W)| \bm W} \\&= \log\bk{ |\bm Z|E_{\tilde Z}\bk{p(T\uppr{k} | \tilde Z\uppr{k},\bm W)p(\tilde Z\uppr{k} | \bm W)} \tag{where $\tilde Z$ is uniform over $\bm Z$}} \\
&= \log E_{\tilde Z}\bk{p(T\uppr{k} | \tilde Z\uppr{k},\bm W)p(\tilde Z\uppr{k} | \bm W)} + \text{const.} \\
&\ge E_{\tilde Z}\bk{\log p(T\uppr{k} | \tilde Z\uppr{k},\bm W)} + E_{\tilde Z}\bk{\log p(\tilde Z\uppr{k} | \bm W)} \tag{Jensen; omitting constant}\\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L \bk{-\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \log \sum_{\bm Z} \exp\pr{-\sum_{Z\uppr{k}} w_i}}  \tag{We parameterize the distribution of $Z\uppr{k}$ as Categorical with $\frac{\exp(-\sum w)}{\sum \exp(-\sum w)}$} \\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L \sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \sum_{i=1}^L \log\bk{|\bm Z| E_{\tilde Z}\bk{\exp\pr{-\sum_{\tilde Z} w_i}}} \\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \sum_{i=1}^L \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i}
\end{align*}

\end{document}

### 2017-11-26 notes_jc.tex
implemented model based on MC

@@
 &= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L \bk{-\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \log \sum_{\bm Z} \exp\pr{-\sum_{Z\uppr{k}} w_i}}  \tag{We parameterize the distribution of $Z\uppr{k}$ as Categorical with $\frac{\exp(-\sum w)}{\sum \exp(-\sum w)}$} \\
-&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L \sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \sum_{i=1}^L \log\bk{|\bm Z| E_{\tilde Z}\bk{\exp\pr{-\sum_{\tilde Z} w_i}}} \\
-&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \sum_{i=1}^L \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i}
+&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L \sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \frac{1}{L}\sum_{i=1}^L \log\bk{|\bm Z| E_{\tilde Z}\bk{\exp\pr{-\sum_{\tilde Z} w_i}}} \\
+&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i}
 \end{align*}

### 2017-12-03 reports/abstract/project-abstract.tex
Clean up repo.

%        File: project-proposal.tex
%     Created: Wed Oct 11 11:00 PM 2017 E
% Last Change: Wed Oct 11 11:00 PM 2017 E
%
\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{tikz}

\pagestyle{fancy}

\lhead{CS 281}
\rhead{Final Project Abstract}

\begin{document}

\title{Final Project Proposal}
\author{
Kevin Chen              \\ \texttt{jiafengchen@college.harvard.edu}
\and Yufeng Ling        \\ \texttt{yufengling@college.harvard.edu}
\and Francisco Rivera   \\ \texttt{frivera@college.harvard.edu}}
\date{\today}

\maketitle

\begin{abstract}
Here.
\end{abstract}

\section{Requirements.txt}

\begin{enumerate}
\item Abstract
\item Changes since proposal
\item Baseline results
\end{enumerate}

\end{document}

### 2017-12-03 reports/proposal/project-proposal.tex
Clean up repo.

%        File: project-proposal.tex
%     Created: Wed Oct 11 11:00 PM 2017 E
% Last Change: Wed Oct 11 11:00 PM 2017 E
%
\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{tikz}

\pagestyle{fancy}

\lhead{CS 281}
\rhead{Final Project Proposal}

\begin{document}

\title{Final Project Proposal}
\author{
Kevin Chen              \\ \texttt{jiafengchen@college.harvard.edu}
\and Yufeng Ling        \\ \texttt{yufengling@college.harvard.edu}
\and Francisco Rivera   \\ \texttt{frivera@college.harvard.edu}}
\date{\today}

\maketitle

\section{Problem Description}

The primary objective of this project is to predict trip duration from taxi data
published by the TLC in New York City. An accurate predictor has a host of
implications, among them,

\begin{itemize}
\item Accurate predictions can allow NYC residents to give themselves a smaller
time buffer when commuting in the city, and thus save time.

\item Estimates of trip duration play a central role in the pricing of trips by
ride-sharing apps, which increasingly feel pressured to a quote a price before
the completion of the trip.

\item From the perspective of a city planner, easing congestion requires an
understanding of which trips systematically take longer.
\end{itemize}

\section{The Data}

\emph{Talk about the data that we get here. Columns we observe, time period we
get, and how much data there is.}

\section{Methodology}

\subsection{Latent Variables Model}

\begin{center}
\begin{tikzpicture}
\draw[->] (0,0) -- (0,5);
\draw[<-] (2,0) -- (2,5);
\draw[->] (4,0) -- (4,5);
\draw[->] (0,0) -- (5,0);
\draw[<-] (0,2) -- (5,2);
\draw[->] (0,4) -- (5,4);

\node[below left] at (0,0) {$(0,0)$};
\node[left] at (0,1) {$\lambda_{0,0}^{\uparrow}$};
\node[left] at (0,1) {$\lambda_{0,0}^{\uparrow}$};
\node[below] at (1,0) {$\lambda_{0,0}^{\rightarrow}$};
\end{tikzpicture}
\end{center}

\end{document}

### 2017-12-03 reports/abstract/abstract.tex

%        File: project-proposal.tex
%     Created: Wed Oct 11 11:00 PM 2017 E
% Last Change: Wed Oct 11 11:00 PM 2017 E
%
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr,amsmath}
\usepackage{macros}
\usepackage{tikz}
\usepackage{palatino, mathpazo, inconsolata}
% \usepackage{natbib}

\pagestyle{fancy}

\lhead{CS 281}
\rhead{Final Project Abstract}

\begin{document}

\title{Abstract and Status Report}
\author{
Jiafeng Chen              \\ \texttt{jiafengchen@college.harvard.edu}
\and Yufeng Ling        \\ \texttt{yufengling@college.harvard.edu}
\and Francisco Rivera   \\ \texttt{frivera@college.harvard.edu}}
\date{\today}

\maketitle

% \section{Requirements.txt}

% \begin{enumerate}
% \item Abstract
% \item Changes since proposal
% \item Baseline results
% \end{enumerate}

\section{Draft abstract}

Using the New York City Taxi and Limousine Commission (TLC) database of taxi rides, we develop a probabilistic model of transportation in New York City. By discretizing the Manhattan geography into a grid structure and by using a mixture model with the latent variable being the route chosen by the driver, we devise a model with a parsimonious number of parameters that is easy to implement even for a large city, fast to estimate, and yields reasonable prediction results relative to benchmarks of linear regression and neural network. Moreover, the parameters of the model can be directly interpreted as measures of congestion on Manhattan's road network.
% We wish to (1) predict the trip duration from a parsimonious set of feature vectors (such as pick-up
% and drop-off location, day of the week and time of the day), and (2) infer variables that represent road condition at a granular level.
% Learning these variables has important practical applications, from helping GPS guidance applications
% choose better routes to helping urban planners and city governments target inefficiencies in urban infrastructure.
% We first established the baseline models using a ridge regression and a neural network, which produced similar results. Next, we construct the exact Manhattan street network, after which we realized the intractability of parameter optimization and opted for a discretized grid of Manhhattan that does not have one-to-one correspondence with the real streets. We then applied a mixture model to the graph with certain assumptions on the ability of the drivers to pick the shortest-time route, which performed better than the baseline model in terms of mean prediction error on the test.

\section{Updates to proposal}

\subsection{Model}

The intuition of our main approach remains the same: break up a trip into its constituent ``sub-trips" and learn the latent variables that correspond to these sub-components. However, in the initial proposal, we aimed to make the level of discretization to be the traversal of an actual street. Using the NYC Street Centerline (CSCL) data from NYC Open Data, we (laboriously) constructed the Manhattan street network, which contained 10296 streets (counting two-way streets as two streets since they will have different parameters in the model; See Figure~\ref{manhattan}).
\begin{figure}
\centering
    \includegraphics[width=0.7\linewidth]{network.png}
    \caption{Complexity of the Manhattan traffic structure as a graph}
    \label{manhattan}
\end{figure}
Given the complexity of the data---both in the number of parameters we need to fit, as well running graph algorithms on a large network---we have since refined our approach. We now aim to discretize Manhattan into a grid and have latent variables for the congestion within each of these discrete units. (See Section~\ref{results})

\subsection{Evaluation}

In our initial proposal, we suggested an $\ell^2$ norm for penalizing predictions would be natural. Upon further inspection of the data, however, we have realized the distribution of travel times has a very heavy right-tail (some trips apparently taking days). Since these observations are either a product of un-clean data or values which we don't particularly care to predict, we opt for reporting percentiles of the $\ell^2$ norm.

\section{Baseline results}

We have two models as baseline. We use a ridge regression and a neural network on the following set of features: trip distance, day of the week, and hour of the day. For the ridge regression, we observed the estimated coefficients of Wednesday to Friday to be the highest among days of the week, and coefficients of 2am-6am to be significantly lower than other hours of the day.
\begin{center}
\begin{tabular}{ccc}\toprule 
    Test-set error descriptive statistics (minutes) & Ridge regression & Neural network \\\midrule
    Mean & 3.684 & 3.544\\
    Standard deviation & 22.177 & 21.786\\
    99\% percentile & 15.268 & 15.059\\\bottomrule
\end{tabular}
\end{center}
Although the neural network model has marginally better performance than the ridge regression and can arguably improve with further parameter tuning, the coefficients of the ridge regression is much more interpretable and generates useful insights. Our approach in Section~\ref{results} shares this feature with the linear regression in generating interpretable coefficients of interest.

\section{Preliminary results}
\label{results}
We implemented a model based on discretization. Since the discretization of Manhattan into literal streets is too cumbersome, we opt to discretize Manhattan into a $P\times Q$ grid (We believe that $40\times 140$ performs well). To put in perspective, each grid point corresponds to roughly a $200\text{ft} \times 200\text{ft}$ area. We convert the coordinates of start and end locations of each trip to the nearest grid point. Let each grid point be associated with a weight $w_i$ whose collection is $\bm W$. These weights represent expected trip times ``passing through'' a grid point.\footnote{If we discretize Manhattan into actual streets, then each street segment would have a weight that indicates how long it takes for a driver to traverse this street segment. Here, the weight of each grid point gives a general sense of how congested a small patch of land is.}

Let $G = \{1,\ldots,N\}$ be the set of grid points. Let $\mathcal{Z}_{ij} \subset 2^G$ be the set of paths from $i$ to $j$.
For each trip $k$ starting from $i$ and ending at $j$, let $T_k$ be the time the trip takes. Let $z_k \in \mathcal{Z}_{ij}$ be a latent variable of the path taken by the trip. We have a mixture model 
\begin{align*}
z_k | \bm W &\sim \mathcal{L}_{\mathcal{Z}_{ij}}\\
T_k | z_k &\sim \mathcal{N}\left(\sum_{i\in z_k} w_i, \sigma^2\right),
\end{align*}
where $\mathcal{L}$ is some distribution over $\mathcal{Z}_{ij}$ and $\bm W = (w_i)_i$ are the weights associated with node $i$---which should be an estimate of time it takes to traverse the grid node.

We assume $\mathcal{L}$ is the following. The driver would only traverse in the direction of her destination. Given hyperparameter $q \in [0,1]$, the driver traverses the shortest-time route ($\argmin_{z} \sum_{z} w_i$) with probability $q$ and randomly picks a route in the direction of her destination with probability $1-q$.

We took an approach that is similar to expectation-maximization and to stochastic variational inference. For each observation, we sample from $\mathcal L_{\mathcal Z_{ij}}$ conditional on the \emph{current} estimates of $\bm W$. We then update the parameters $\bm W$ via the gradients of the squared error loss between the predicted trip time and the actual trip time (thus corresponding to a Normal model with known variance). This approach is similar to \cite{zhan2013urban}, who also use a mixture model, while using a discrete choice model for modeling $\mathcal L_{\mathcal Z_{ij}}$. However, the model in \cite{zhan2013urban} runs on a much smaller scale facing larger computational constraints, whereas we are able to estimate on a larger scale by using a sampling approach for $\mathcal L_{\mathcal Z_{ij}}$.

There are a few interesting results to note:
\begin{enumerate}
    \item The model performs well even with a small dimension of points. In Figure~\ref{errs}, we use a model with only 350 parameters (about 1/5 of the neutral network), we are able to achieve test-set error rates that are on par with the neural net. Moreover, we have not stratified the data by time and week, which are important predictors of trip time in the model.
    \item The visualization of the model is telling. In all three dimensions of the model that we tested (Figure~\ref{small} and Figure~\ref{larger}), the visualization of the model parameters are able to identify similar areas of high and low congestion. This is robust to dimension, initialization, and values of $q$ (Figure~\ref{nosmooth}). 
    \item The model is a bit sensitive to extreme values (which we suspect to be data errors). The extreme values tend to be unusually long trips, and they tend to update the parameters in extreme directions (the red lines in Figures~\ref{larger} and \ref{nosmooth}). Though these errors fade out over time (see attached 100-frame video of visualizations with one frame for 2000 trips\footnote{The images are snapshots of parameter values in the updating process}), they fade out especially slowly since trips that take long are unlikely to be selected if $q$ is high, and thus it's hard for high parameter values to be corrected on later conditioning.
    \item We also observe that $q$ acts as a smoothing parameter. With $q=1$ in Figure~\ref{nosmooth}, the parameters rigid, forming clear lines that probably represents the disproportionate effects of only a few trips.
\end{enumerate}

\subsection{Future directions}
Here are a few areas we are looking to explore:
\begin{itemize}
    \item We are looking to explore the theoretical properties of the model and estimation approach by formalizing them in terms of EM and stochastic variational inference.
    \item We will split the training sample by some observable characteristics such as day of the week and time of day, and train on each subsample. This approach should improve prediction accuracy, and provide more insight in the evolution of congestion in the temporal dimension.
    \item We will consider remedies of extreme values. Currently, since we assume drivers behave intelligently, weights that are overly large (usually due to unusually long trips because of unclean data) tend to very slowly be corrected, since intelligent drivers avoid such routes and prevents the routes from being observed and updated. Graphically, this results in abrupt red lines across our Manhattan map and only updated gradually.
    \item We will attempt to address the intractability of parameter optimization on the exact street network by dividing Manhattan into neighborhoods and optimize the street weights by training on a subsample of the dataset restricted to each neighborhood. With the weight of each streets calculated, we then make prediction on the complete test set.
\end{itemize}

\bibliographystyle{IEEEtran}
\bibliography{refs}

\appendix
\section{Attached figures}

\begin{figure}[!h]
\centering
\includegraphics[width=0.7\textwidth]{error.png}
\caption{The distribution of error obtained on the test set by using a low-dimensional model with $P=10, Q = 35$ and $q=1$. The mean loss is $3.28$ minutes and the 99\% is $15.49$ minutes in error.}
\label{errs}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.7\textwidth]{low-dim1.png}
\caption{We plot via \texttt{plt.imshow(...,cmap='hot')} the distribution of parameter values since they have physical interpretations on a heat map. The right panel is the distribution of trip starting locations which trace out the shape of Manhattan. This is on a low dimensional model with $P=10, Q = 35$. Here, brighter colors (yellow) means higher parameter value, which points to more congestion. We also initialize grid points that are above non-terrestrial areas (e.g. Hudson river) to have high values and omit them in the plotting. We later realized that the initialization does not matter, and the model is able to pick up the shape of Manhattan automatically.}
\label{small}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.4\textwidth]{91.png}
\includegraphics[width=0.39\textwidth]{large.png}
\caption{Left: an iteration of the model with $40 \times 140$ parameters. Right: an iteration of the model with $80\times 280$ parameters. In both, $q=0.5$.}
\label{larger}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.4\textwidth]{no_smooth.png}
\caption{An iteration of the model with $q = 1$. We see that without smoothing, the parameters doesn't really look right. }
\label{nosmooth}
\end{figure}

\end{document}

### 2017-12-03 reports/paper/paper.tex
.

\documentclass{article}

\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{lipsum}
\usepackage{todonotes}
\usepackage{jchen2}
\usepackage[accepted]{icml2017}
\icmltitlerunning{CS 281 Final Project Template}

\begin{document}

\twocolumn[
\icmltitle{Project Title}
\begin{icmlauthorlist}
  \icmlauthor{Jiafeng Chen}{}
\icmlauthor{Yufeng Ling}{}
\icmlauthor{Francisco Rivera}{}
\end{icmlauthorlist}

\vskip 0.3in
]

\begin{abstract}
  \begin{itemize}
  \item This document describes the expected style, structure, and rough proportions for your final project write-up.
  \item While you are free to break from this structure, consider it a strong prior for our expectations of the final report.
\item Length is a hard constraint. You are only allowed max \textbf{8 pages} in this format. While you can include supplementary material, it will not be factored into the grading process. It is your responsibility to convey the main contributions of the work in the length given.
  \end{itemize}

\end{abstract}

\section{Introduction}
\label{sec:introduction}

Example Structure:
\begin{itemize}
\item What is the problem of interest and what (high-level) are the current best methods for solving it?
\item How do you plan to improve/understand/modify this or related methods?
\item Preview your research process, list the contributions you made, and summarize your experimental findings.
\end{itemize}

\section{Background}
Example Structure:
\begin{itemize}
\item What information does a non-expert need to know about the problem domain?
\item What data exists for this problem?
\item What are the challenges/opportunities inherent to the data? (High dimensional, sparse, missing data, noise, structure, discrete/continuous, etc?)
\end{itemize}

\section{Related Work}

Example Structure:
\begin{itemize}
\item What 3-5 papers have been published in this space?
\item How do these  differ from your approach?
\item What data or methodologies do each of these works use?
\item How do you plan to compare to these methods?
\end{itemize}

\section{Model}

Example Structure:

\begin{itemize}
\item What is the formal definition of your problem?
\item What is the precise mathematical model you are using to represent it? In almost all cases this will use the probabilistic language from class, e.g.
  \begin{equation}
  z \sim {\cal N}(0, \sigma^2)\label{eq:1}
\end{equation}
But it may also be a neural network, or a non-probabilistic loss,
\[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]

This is also a good place to reference a diagram such as Figure~\ref{fig:diagram}.

\item What are the parameters or latent variables of this model that you plan on estimating or inferring? Be explicit. How many are there? Which are you assuming are given? How do these relate to the original problem description?
\end{itemize}

% \begin{figure}
%   \centering
%   \missingfigure[figheight=8cm]{}
%   \caption{\label{fig:diagram} This is a good place to include a diagram showing how your model works. Examples include a graphical model or a neural network block diagram.}
% \end{figure}

\section{Inference (or Training)}

\begin{itemize}
\item How do you plan on training your parameters / inferring the
  states of your latent variables (MLE / MAP / Backprop / VI / EM / BP / ...)

\item What are the assumptions implicit in this technique? Is it an approximation or exact? If it is an approximation what bound does it optimize?

\item What is the explicit method / algorithm that you derive for learning these parameters?
\end{itemize}

\begin{algorithm}
  \begin{algorithmic}
    \STATE{}
  \end{algorithmic}
  \caption{Your Pseudocode}
\end{algorithm}

\section{Methods}

\begin{itemize}
\item What are the exact details of the dataset that you used? (Number of data points / standard or non-standard / synthetic or real / exact form of the data)

\item What are the exact details of the features you computed?

\item How did you train or run inference? (Optimization method / hyperparameter settings / amount of time ran / what did you implement versus borrow / how were baselines computed).

\item What are the exact details of the metric used?
\end{itemize}

\section{Results}

\begin{itemize}
\item What were the results comparing previous work / baseline systems / your systems on the main task?
\item What were the secondary results comparing the variants of your system?
\item This section should be fact based and relatively dry. What happened, what was significant?
\end{itemize}

\begin{table*}
  \centering
  \missingfigure{}
  \caption{This is usually a table. Tables with numbers are generally easier to read than graphs, so prefer when possible.}
  \label{fig:mainres}
\end{table*}

% \begin{table}
%   \centering
%   \missingfigure[figheight=5cm]{}
%   \caption{Secondary table or figure in results section.}
%   \label{fig:mainres}
% \end{table}

\section{Discussion}

\begin{itemize}
\item What conclusions can you draw from the results section?
\item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.

\item  What questions remain open? What did you think might work, but did not?
\end{itemize}

% \begin{figure}
%   \centering
%   \missingfigure{}
%   \missingfigure{}
%   \missingfigure{}
%   \caption{Visualizations of the internals of the system.}
% \end{figure}

\section{Conclusion}

\begin{itemize}
\item What happened?
\item What next?
\end{itemize}

% \section*{Acknowledgements}

% \textbf{Do not} include acknowledgements in the initial version of
% the paper submitted for blind review.

% If a paper is accepted, the final camera-ready version can (and
% probably should) include acknowledgements. In this case, please
% place such acknowledgements in an unnumbered section at the
% end of the paper. Typically, this will include thanks to reviewers
% who gave useful comments, to colleagues who contributed to the ideas,
% and to funding agencies and corporate sponsors that provided financial
% support.

\bibliography{example}
\bibliographystyle{icml2017}

\end{document}

### 2017-12-04 reports/paper/paper.tex
.

@@
 \usepackage{lipsum}
-\usepackage{todonotes}
+\usepackage{todonotes,tikz}
+\usetikzlibrary{bayesnet}
 \usepackage{jchen2}
 \usepackage[accepted]{icml2017}
-\icmltitlerunning{CS 281 Final Project Template}
+\icmltitlerunning{Project Title}
 
@@
 
+We represent a city's road network with a connected graph $G = (V,E)$.
+Assume that each vertex $i \in V$ is associated with a weight $w_i$, representing the cost of traversing vertex $i$.
+A trip is represented by a path in $G$, and the distribution of the trip's duration depends on the weights $w_i$ of vertices included in the path.
+Note that the choice of the path can in general depend on the collection of weights $\bm W$.
+In full generality, the model is represented by Figure~\ref{fig:dgm}, where trips in the data are indexed by $(k)$, $T\uppr{k}$ is the observed trip duration, and $Z\uppr{k}$ is the path taken by trip $k$, a latent variable.
+
+\begin{figure}[h]
+  \centering
+  \caption{Representation of model as a directed graph}
+  \label{fig:dgm}
+  \vspace{1em}
+  \begin{tikzpicture}
+    \node[latent] (W) {$\bm W$};
+    \node[latent, below=of W] (Z) {$Z\uppr{k}$};
+    \node[latent, right=of Z] (T) {$T\uppr{k}$};
+    \edge{W} {Z};
+    \edge{W, Z} {T};
+    \plate[inner sep=0.5em, yshift=0.2em] {ZT} {(Z) (T)} {$k=1,\ldots,N$} ;
+  \end{tikzpicture}
+\end{figure}
+
+\subsection{Parameterization}
+
+In principle, in our application to the New York City taxi data, we may take $G$ to be a graph representing the exact road network in New York City, where each vertex is a \emph{road segment} and the directed edge $(i,j) \in E$ if one can drive directly onto road segment $j$ from road segment $i$; such a parameterization allows $w_i$ to be directly interpretable as a measure of congestion on road segment $i$.
+However, such a detailed construction presents serious computational challenges when training on a large dataset, since solving path-finding problems and computing minimal paths are nontrivially expensive.\footnote{Manhattan has on the order of $10^4$ road segments, and the dataset contains the order of $10^7$ trips for January 2009 alone.}
+
+To avoid these challenges, we parameterize $G$ as an undirected rectangular grid.
+Despite not being able to pinpoint weights $w_i$ to congestion of specific road segments, we are nonetheless able to interpret the weights $w_i$ as representative of congestion on a small patch of land.
+We may now represent a path $Z\uppr{k}$ as a set of indices $i$ of grid points traversed by the path.
+In full generality, there are an infinite number of paths connecting any two points $i,j$ on the grid, but the vast majority of these paths are not sensible.
+Thus we restrict the set of possible paths for trip $k$ to a \emph{set of reasonable paths} $\bm Z\uppr{k}$, where each path in $\bm Z\uppr{k}$ travels strictly in the direction of the destination.
+For instance, if the destination of $j$ is to the northeast of the starting location $i$, then the set of reasonable paths $\bm Z$ are the set of paths that only involve northward or eastward movements (e.g.~Figure~\ref{fig:reasonable}).
+
+\begin{figure}[h]
+\centering
+\caption{An example of a reasonable path}
+\vspace{1em}
+
+\label{fig:reasonable}
+\begin{tikzpicture}
+\draw (-0.5,0) -- (3.5,0);
+\draw (-0.5,1) -- (3.5,1);
+\draw (-0.5,2) -- (3.5,2);
+
+\draw (0,-0.5) -- (0,2.5);
+\draw (1,-0.5) -- (1,2.5);
+\draw (2,-0.5) -- (2,2.5);
+\draw (3,-0.5) -- (3,2.5);
+
+\foreach \y in {0,1}
+{
+  \foreach \x in {0,...,2}
+  {
+    \node at (\x+.5,\y+.5) {$w_{\x,\y}$};
+  }
+}
+
+\draw[line width = 12pt, opacity=.2] (0.2,.5) -- (1.5,.5) -- (1.5,1.5) -- (2.8,1.5);
+\end{tikzpicture}
+\end{figure}
+
+We have the following model: \begin{align*}
+\bm W &\sim p(\bm W) \\
+Z\uppr{k} &\sim p(Z\uppr{k} | \bm W) \\
+T\uppr{k} | \bm W, Z\uppr{k} &\sim \Norm\pr{\sum_{i\in Z\uppr{k}} w_i, \sigma^2},
+\end{align*}
+
+{
+\color{gray}
 Example Structure:
@@
 \end{itemize}
+}
 

### 2017-12-08 reports/paper/paper.tex
.

@@
 \usepackage{todonotes,tikz}
+\usepackage{cite}
+\usepackage{natbib}
+
 \usetikzlibrary{bayesnet}
@@
 In full generality, the model is represented by Figure~\ref{fig:dgm}, where trips in the data are indexed by $(k)$, $T\uppr{k}$ is the observed trip duration, and $Z\uppr{k}$ is the path taken by trip $k$, a latent variable.
+Our primary interest is to perform inference on $\bm W$, so as to learn the levels of congestion associated with each vertex in $G$.
 
@@
 Thus we restrict the set of possible paths for trip $k$ to a \emph{set of reasonable paths} $\bm Z\uppr{k}$, where each path in $\bm Z\uppr{k}$ travels strictly in the direction of the destination.
-For instance, if the destination of $j$ is to the northeast of the starting location $i$, then the set of reasonable paths $\bm Z$ are the set of paths that only involve northward or eastward movements (e.g.~Figure~\ref{fig:reasonable}).
+For instance, if the destination of $j$ is to the northeast of the starting location $i$, then the set of reasonable paths $\bm Z$ are the set of paths that only involve northward or eastward movements (e.g.~Figure~\ref{fig:reasonable} shows a reasonable path from $(0,0)$ to $(2,1)$).
+Such a parameterization is more general than many in the literature; \citet{zhan2013urban}, for instance, uses the $K$-shortest path algorithm and considers the shortest 20 paths as a set of reasonable paths.
 
@@
 
-We have the following model: \begin{align*}
+We parameterize the conditional distribution of $T\uppr{k}$ as Normal, in the following reformulation of the directed graphical model: \begin{align*}
 \bm W &\sim p(\bm W) \\
@@
 \end{align*}
+where $p(Z\uppr{k} | \bm W)$ is a distribution over $\bm Z^{\uppr{k}}$.
+We consider two different ways to parameterize $p(Z\uppr{k} | \bm W)$: softmax regression and uniform.
+In the \emph{softmax regression} model, a type of generalized linear model for discrete choice problems \cite{mcfadden1973conditional}, we parameterize the route choice such that \[
+p(Z\uppr{k}|\bm W) \propto \exp\pr{-\sum_{i\in Z\uppr{k}}w_i},
+\]
+in order to encode the fact that drivers avoid routes that take a long period of time.
+In the \emph{uniform} model, we simply assume that route choice is independent and uniform on the set of reasonable paths: \[
+p(Z\uppr{k}|\bm W) \propto 1.
+\]
+The uniform model trades off realism in modeling for improvement in computation and training, as we see in Section~\ref{sec:inference}.
+
+In our application to the Manhattan dataset, we perform MLE inference, or, equivalently, MAP inference with $p(\bm W) \propto 1$.
+In principle, it is not difficult to parameterize the prior of $\bm W$ as an undirected graphical model, since we need only to supply edge and unary potentials.
+For instance, to impose a correlated prior on $\bm W$, as suggested by some \cite{hunter2009path}, we simply penalize large differences in neighboring weights in the edge potential, effectively assuming a prior model that is similar to a continuous version of the Ising model.
 
@@
 \section{Inference (or Training)}
+\label{sec:inference}
+We perform maximum likelihood inference, maximizing \[
+\max_{\bm W} \log p(\{T\uppr{k}\}_{k=1}^N|\bm W) = \max_{\bm W} \sum_{k=1}^N \log p(T\uppr{k} | \bm W).
+\]
+The log-likelihood is 
+\begin{align*}
+&\log p (T\uppr{k} | \bm W) \\
+=& \log\pr{\sum_{Z\uppr{k} \in \bm Z\uppr{k}} p(T\uppr{k}|Z\uppr{k},\bm W)p(Z\uppr{k}|\bm W)} \\
+=& \log\pr{\E_{Z\uppr{k}}\bk{p(T\uppr{k}|Z\uppr{k},\bm W)}}.
+\end{align*}
+The expectation is a sum of the size $|\bm Z\uppr{k}|,$ which, for an $m\times n$ trip\footnote{By an $m\times n$ trip, we mean a trip with east-west distance $n$ and north-south distance $m$}, is of $\binom{m+n}{n} \approx \frac{(n+m)^n}{n^n}e^n$ terms.
+Computing this expectation is the main inference challenge of our project.
 
+\subsection{Inference in the uniform model}
+
+By assuming the uniform model $p(Z|\bm W) \propto 1$, we gain the ability to work with an expectation over $\bm W$ instead of an expectation over $\bm Z$, since the probability that a particular weight is included in a path is readily computable from elementary combinatorics.
+We maximize an approximate lower bound of the log-likelihood by applications of Jensen's inequality: \begin{align*}
+\ell(\bm W) &= \log\pr{\E_{Z\uppr{k}}\bk{p(T\uppr{k}|Z\uppr{k},\bm W)}} \\
+&\ge -\frac{1}{2\sigma^2}\pr{T\uppr{k} - \E\bk{\sum_{i\in Z\uppr{k}} w_i}}^2 + \text{const.} \tag{$*$} \\
+&= -\frac{1}{2\sigma^2}\pr{T\uppr{k} - \sum_{i} w_i\pi_i}^2 + \text{const.},
+\end{align*}
+where $\pi_i$ is the marginal probability of node $i$ being included in a uniform route.\footnote{$\pi_i$ can be computed analytically.
+Suppose the source and destination of the trip are $(n,m)$ apart and vertex $i$ is $(a,b)$ away from the source.
+Then, by elementary combinatorics, \[\pi_i = \frac{\binom{a+b}{a}\binom{n+m-a-b}{n-a}}{\binom{n+m}{n}}\]} Note that $(*)$ is only a lower bound if $T\uppr{k} - \sum w_i$ is small compared to $\sigma^2$, since for small values of $T\uppr{k} - \sum w_i$ relative to $\sigma$, the Normal density is concave, and the inequality follows by Jensen's inequality.
+We indeed make this assumption.
+By assuming a structure of $Z|\bm W$, we gain a great deal of convenience in computation, effectively reducing a sum of $\binom{n+m}{n}$ terms to a sum of merely $nm$ terms.
+
+\subsection{Inference in the softmax regression model}
+
+{\color{gray}
 \begin{itemize}
@@
 \end{algorithm}
+}
 
@@
 
-\bibliography{example}
+\bibliography{biblio.bib}
 \bibliographystyle{icml2017}

### 2017-12-08 reports/paper/paper.tex
Working on Background section of paper.

@@
 \usepackage[accepted]{icml2017}
-\icmltitlerunning{Project Title}
+\icmltitlerunning{Probabilistic Inference of NYC Congestion from Taxi Data}
 
@@
 \twocolumn[
-\icmltitle{Project Title}
+\icmltitle{Probabilistic Inference of NYC Congestion from Taxi Data}
 \begin{icmlauthorlist}
@@
 \section{Background}
+{
+\color{gray}
 Example Structure:
@@
 \end{itemize}
+}
+
+The NYC Taxi and Limousine Commission (TLC) makes
+available\footnote{\url{http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml}}
+a dataset of taxi rides taken in the city. For our purposes, each ride is
+characterized by the time and location (latitude and longitude) of the pickup as
+well as the corresponding quantities for the drop-off\footnote{Additional
+information orthogonal to predicting trip duration---such as passenger
+counts and tip---is discarded}. Moreover, there is an abundance of data. In
+January of 2009 alone, there were over 14 million taxi rides recorded.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{figs/ridesbyhour.pdf}
+\caption{Taxi Trips by Time of Day}
+\label{fig:ridesbyhour}
+\end{figure}
+
+When we aggregate by hour of day (Figure \ref{fig:ridesbyhour}), we see some variance with anywhere between
+a hundred thousand and a million trips per interval. This is reassuring for
+future bucketing of the data.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{figs/tripdurations.pdf}
+\caption{Taxi Trips by Duration}
+\label{fig:tripdurations}
+\end{figure}
+
+While we are not lacking in data volume, two important challenges do appear when
+working with this dataset. The first is that the data set is not completely clean.
+Case in point, while most trip durations (Figure \ref{fig:tripdurations}) are
+reasonable for a taxi trip, just over 1\% of trips have a drop-off time that is
+before the pick-up time, making for a trip with negative duration. Furthermore,
+the longest trip has duration of just over 42 days. It is difficult to
+accurately assess
+the veracity of the pickup and dropoff coordinates, but we observe
+that some trips
+have geographic coordinates that place the start or end of the trip in the
+ocean.
+
+Additionally, whereas we would like to infer road conditions along the route of
+a taxi trip, we can only observe the start and end-points. The fact that the
+route taken is unobservable is primarily responsible for our inference
+challenges.
 
@@
 
-This is also a good place to reference a diagram such as Figure~\ref{fig:diagram}.
+% Commented out dangling reference --FR
+This is also a good place to reference a diagram such as %Figure~\ref{fig:diagram}.
 
@@
 \ell(\bm W) &= \log\pr{\E_{Z\uppr{k}}\bk{p(T\uppr{k}|Z\uppr{k},\bm W)}} \\
-&\ge -\frac{1}{2\sigma^2}\pr{T\uppr{k} - \E\bk{\sum_{i\in Z\uppr{k}} w_i}}^2 + \text{const.} \tag{$*$} \\
-&= -\frac{1}{2\sigma^2}\pr{T\uppr{k} - \sum_{i} w_i\pi_i}^2 + \text{const.},
+&= \log\pr{
+\E_{Z\uppr{k}}\bk{
+\frac{1}{\sigma\sqrt{2\pi}} e^{
+-\frac{1}{2\sigma^2}(T\uppr{k} - \sum w_i)^2
+}
+}
+} 
+% p(T\uppr{k}|Z\uppr{k},\bm W)}}
 \end{align*}
-where $\pi_i$ is the marginal probability of node $i$ being included in a uniform route.\footnote{$\pi_i$ can be computed analytically.
+Note that the Normal density is concave for $|T\uppr{k} - \sum w_i| < \sigma$ and convex otherwise. Applying Jensen's inequality locally, we obtain that \[
+B(\bm W) = \frac{1}{2\sigma^2}\pr{T\uppr{k} - \sum_i w_i \pi_i}^2,
+\]
+where $\pi_i$ is the marginal probability of node $i$ being included in a uniform route,\footnote{$\pi_i$ can be computed analytically.
+Suppose the source and destination of the trip are $(n,m)$ apart and vertex $i$ is $(a,b)$ away from the source.
+Then, by elementary combinatorics, \[\pi_i = \frac{\binom{a+b}{a}\binom{n+m-a-b}{n-a}}{\binom{n+m}{n}}\]} is a local lower bound for the negative log-likelihood for trips that we predict poorly and is a local upper bound for trips that we predict well, up to a constant.
+Thus $B(\bm W)$ is a good approximation of the objective function that is easily computable, involving only $mn$ as opposed to $\binom{m+n}{n}$ terms.
+
+% &\ge -\frac{1}{2\sigma^2}\pr{T\uppr{k} - \E\bk{\sum_{i\in Z\uppr{k}} w_i}}^2 + \text{const.} \tag{$*$} \\
+% &= -\frac{1}{2\sigma^2}\pr{T\uppr{k} - \sum_{i} w_i\pi_i}^2 + \text{const.},
+% \end{align*}
+% where $\pi_i$ is the marginal probability of node $i$ being included in a uniform route.\footnote{$\pi_i$ can be computed analytically.
 Suppose the source and destination of the trip are $(n,m)$ apart and vertex $i$ is $(a,b)$ away from the source.
@@
 \subsection{Inference in the softmax regression model}
+
+We now consider a more realistic but more complex model, where we parameterize $Z|\bm W$ as a GLM, namely as a softmax regression where $p(Z|\bm W) \propto \exp\pr{-\sum_{i\in Z}w_i}$, encoding drivers' preferences for shorter trips.
+The log-likelihood in this model is \begin{align*}
+& \ell(\bm W) \\
+=& \log\pr{\E_{Z\uppr{k}}\bk{\frac{1}{\sqrt{2\pi}\sigma} e^{-\frac{1}{2\sigma^2}(T\uppr{k} - \sum_i w_i)^2}}}\\
+=& \log\pr{\sum_{Z\uppr{k}\in \bm Z\uppr{k}} \frac{1}{\sqrt{2\pi}\sigma} e^{-\frac{1}{2\sigma^2}(T\uppr{k} - \sum_i w_i)^2} p(Z\uppr{k}|\bm W)}
+\end{align*}
+
+Before we discuss our inference strategy, we first discuss a few difficulties of the model.
+At first glance, the latent variable structure seems lend well to the Expectation-Maximization algorithm \cite{dempster1977maximum}.
+However, the EM algorithm requires computing the term \[
+Q(\bm W|\bm W\uppr{t}) = \E\pr{\log(T\uppr{k}, Z\uppr{k} | \bm W) | T\uppr{k}, \bm W\uppr{t}},
+\]
+and computing the expectation requires the conditional distribution $p(Z | T, \bm W) = \frac{p(T|Z,\bm W)p(Z|\bm W)}{p(T|\bm W)}$, where the denominator is intractable to compute.
+We might also consider techniques in variational inference \cite{blei2017variational}.
+However, the latent variable $Z\uppr{k}$ is a random set of indices with the property that the indices form a path on the grid, and thus cannot be reasonably decomposed into independent components, ruling out the mean-field algorithm.
+Another promising option is stochastic gradient variational Bayes (SGVB) and variational autoencoders \cite{kingma2013auto}, which is designed for large datasets for which the EM algorithm fails.
+However, SGVB requires a reparameterization of the latent variable drawn from an approximate distribution, in order for the gradients to be properly computed.
+In SGVB, one independently draws some $\epsilon$ from a distribution and obtains $z$ via $z = g(\epsilon)$ for some continuous function $g$.\footnote{Here we are using the notation in \cite{kingma2013auto}.} This is difficult to do in our context, since $Z$ is not continuous nor scalar-valued.
+Thus it is difficult to find an $\epsilon$ and a continuous transformation to approximate samples from the distribution of $Z$.
+
+Our inference strategy is based on a sampling-based approximating to the expectation over $\bm Z$. The key trick we use is that 
+\begin{equation}
+\label{eq:trick}
+  \E_{Z}\bk{f(Z\uppr{k})} = |\bm Z\uppr{k}| \E_{\tilde Z}\bk{f(\tilde Z\uppr{k})p_Z(\tilde Z\uppr{k}|\bm W)},
+\end{equation}
+where $\tilde Z$ is drawn uniformly from $\bm Z$.
+This is the same technique as in importance sampling, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations.
+Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields the following lower bound for negative log-likelihood, up to a constant,
+\begin{align*}
+&\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} + \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i \\&\qquad + \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i},
+\end{align*}
+where we sample uniformly and independently $\{\tilde Z_j\uppr{k}\}_{j=1}^L$.
 

### 2017-12-08 reports/paper/paper.tex
model

@@
 \usepackage[accepted]{icml2017}
-\icmltitlerunning{Probabilistic Inference of NYC Congestion from Taxi Data}
+\icmltitlerunning{Project Title}
 
@@
 \twocolumn[
-\icmltitle{Probabilistic Inference of NYC Congestion from Taxi Data}
+\icmltitle{Project Title}
 \begin{icmlauthorlist}
@@
 \section{Background}
-{
-\color{gray}
 Example Structure:
@@
 \end{itemize}
-}
-
-The NYC Taxi and Limousine Commission (TLC) makes
-available\footnote{\url{http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml}}
-a dataset of taxi rides taken in the city. For our purposes, each ride is
-characterized by the time and location (latitude and longitude) of the pickup as
-well as the corresponding quantities for the drop-off\footnote{Additional
-information orthogonal to predicting trip duration---such as passenger
-counts and tip---is discarded}. Moreover, there is an abundance of data. In
-January of 2009 alone, there were over 14 million taxi rides recorded.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/ridesbyhour.pdf}
-\caption{Taxi Trips by Time of Day}
-\label{fig:ridesbyhour}
-\end{figure}
-
-When we aggregate by hour of day (Figure \ref{fig:ridesbyhour}), we see some variance with anywhere between
-a hundred thousand and a million trips per interval. This is reassuring for
-future bucketing of the data.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/tripdurations.pdf}
-\caption{Taxi Trips by Duration}
-\label{fig:tripdurations}
-\end{figure}
-
-While we are not lacking in data volume, two important challenges do appear when
-working with this dataset. The first is that the data set is not completely clean.
-Case in point, while most trip durations (Figure \ref{fig:tripdurations}) are
-reasonable for a taxi trip, just over 1\% of trips have a drop-off time that is
-before the pick-up time, making for a trip with negative duration. Furthermore,
-the longest trip has duration of just over 42 days. It is difficult to
-accurately assess
-the veracity of the pickup and dropoff coordinates, but we observe
-that some trips
-have geographic coordinates that place the start or end of the trip in the
-ocean.
-
-Additionally, whereas we would like to infer road conditions along the route of
-a taxi trip, we can only observe the start and end-points. The fact that the
-route taken is unobservable is primarily responsible for our inference
-challenges.
 
@@
 
-% Commented out dangling reference --FR
-This is also a good place to reference a diagram such as %Figure~\ref{fig:diagram}.
+This is also a good place to reference a diagram such as Figure~\ref{fig:diagram}.
 

### 2017-12-09 reports/paper/paper.tex

--- 2017-12-08 reports/paper/paper.tex
@@
 
+{
+\color{gray}
 \begin{itemize}
@@
 \end{itemize}
+}
+
+For our dataset, we started with the TLC dataset of all NYC taxi trips taken in
+January 2009. From the raw data, several computed columns were added.
+One-hot-encoded columns were added for hour of day (24 columns) as well as for
+day of the week (7 columns). The duration of the trip was also explicitly
+computed in seconds as the dependent variable to predict.
+
+In addition to adding columns, the original dataset was filtered. Geographic
+coordinates were\ldots
 

### 2017-12-09 reports/paper/paper.tex

@@
 \section{Background}
-{
-\color{gray}
-Example Structure:
-\begin{itemize}
-\item What information does a non-expert need to know about the problem domain?
-\item What data exists for this problem?
-\item What are the challenges/opportunities inherent to the data? (High dimensional, sparse, missing data, noise, structure, discrete/continuous, etc?)
-\end{itemize}
-}
+% {
+% \color{gray}
+% Example Structure:
+% \begin{itemize}
+% \item What information does a non-expert need to know about the problem domain?
+% \item What data exists for this problem?
+% \item What are the challenges/opportunities inherent to the data? (High dimensional, sparse, missing data, noise, structure, discrete/continuous, etc?)
+% \end{itemize}
+% }
 
@@
 
-{
-\color{gray}
-Example Structure:
-
-\begin{itemize}
-\item What is the formal definition of your problem?
-\item What is the precise mathematical model you are using to represent it? In almost all cases this will use the probabilistic language from class, e.g.
-  \begin{equation}
-  z \sim {\cal N}(0, \sigma^2)\label{eq:1}
-\end{equation}
-But it may also be a neural network, or a non-probabilistic loss,
-\[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]
-
-% Commented out dangling reference --FR
-This is also a good place to reference a diagram such as %Figure~\ref{fig:diagram}.
-
-\item What are the parameters or latent variables of this model that you plan on estimating or inferring?
+% {
+% \color{gray}
+% Example Structure:
+% 
+% \begin{itemize}
+% \item What is the formal definition of your problem?
+% \item What is the precise mathematical model you are using to represent it? In almost all cases this will use the probabilistic language from class, e.g.
+%   \begin{equation}
+%   z \sim {\cal N}(0, \sigma^2)\label{eq:1}
+% \end{equation}
+% But it may also be a neural network, or a non-probabilistic loss,
+% \[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]
+% 
+% % Commented out dangling reference --FR
+% This is also a good place to reference a diagram such as %Figure~\ref{fig:diagram}.
+% 
+% \item What are the parameters or latent variables of this model that you plan on estimating or inferring?
 Be explicit.
@@
 How do these relate to the original problem description?
-\end{itemize}
-}
+% \end{itemize}
+% }
 
@@
 
-{
-\color{gray}
-\begin{itemize}
-\item What are the exact details of the dataset that you used? (Number of data points / standard or non-standard / synthetic or real / exact form of the data)
-
-\item What are the exact details of the features you computed?
-
-\item How did you train or run inference? (Optimization method / hyperparameter settings / amount of time ran / what did you implement versus borrow / how were baselines computed).
-
-\item What are the exact details of the metric used?
-\end{itemize}
-}
+% {
+% \color{gray}
+% \begin{itemize}
+% \item What are the exact details of the dataset that you used? (Number of data points / standard or non-standard / synthetic or real / exact form of the data)
+% 
+% \item What are the exact details of the features you computed?
+% 
+% \item How did you train or run inference? (Optimization method / hyperparameter settings / amount of time ran / what did you implement versus borrow / how were baselines computed).
+% 
+% \item What are the exact details of the metric used?
+% \end{itemize}
+% }
 

### 2017-12-09 test.py
updated test

@@ -11,9 +11,11 @@ def test(predict, result_file,
 
 
     errs = []
-    for _, duration, sx, sy, tx, ty in tqdm(df.itertuples()):
-        pred = predict(sx, sy, tx, ty)
-        errs.append(duration - pred)
+    with tqdm(total=len(df)) as pbar:
+        for _, duration, sx, sy, tx, ty in (df.itertuples()):
+            pbar.update(1)
+            pred = predict(sx, sy, tx, ty)
+            errs.append(duration - pred)
 
     errs = pd.Series(errs)
 

### 2017-12-09 test.py
,

@@ -17,7 +17,7 @@ def test(predict, result_file,
             pred = predict(sx, sy, tx, ty)
             errs.append(duration - pred)
 
-    errs = pd.Series(errs)
+    errs = pd.Series(errs)/60
 
     stat = {
         'sd_duration' : df['duration'].std(),
@@ -29,7 +29,7 @@ def test(predict, result_file,
         'test_r2' : 1 - np.std(errs) ** 2 / (df['duration'].std() ** 2)
     }
 
-    json.dump(stat, open(result_file, 'w'))
-    pickle.dump(errs, open(result_file + '.raw', 'wb'))
+    json.dump(stat, open(result_file + '.json', 'w'))
+    pickle.dump(errs, open(result_file + '.pickle', 'wb'))
 
     return errs

### 2017-12-09 test.py
sad

@@ -20,7 +20,7 @@ def test(predict, result_file,
     errs = pd.Series(errs)/60
 
     stat = {
-        'sd_duration' : df['duration'].std(),
+        'sd_duration' : df['duration'].std() / 60,
         'mean_err': errs.mean(),
         'sd_err': errs.std(),
         'mean_abs_err': errs.abs().mean(),

### 2017-12-09 test.py
/60

@@ -7,6 +7,7 @@ import pickle
 def test(predict, result_file,
          test_file="data/split_test/JC_week1_hour8_test_40x140.csv"):
     df = pd.read_csv(test_file)
+    df['duration'] = df['duration'] / 60
     df = df.loc[:, ['duration', 'sx', 'sy', 'tx', 'ty']]
 
 
@@ -14,13 +15,13 @@ def test(predict, result_file,
     with tqdm(total=len(df)) as pbar:
         for _, duration, sx, sy, tx, ty in (df.itertuples()):
             pbar.update(1)
-            pred = predict(sx, sy, tx, ty)
+            pred = predict(sx, sy, tx, ty) / 60
             errs.append(duration - pred)
 
-    errs = pd.Series(errs)/60
+    errs = pd.Series(errs)
 
     stat = {
-        'sd_duration' : df['duration'].std() / 60,
+        'sd_duration' : df['duration'].std(),
         'mean_err': errs.mean(),
         'sd_err': errs.std(),
         'mean_abs_err': errs.abs().mean(),

### 2017-12-09 test.py
Latest itration of test.py

@@ -21,7 +21,7 @@ def test(predict, result_file,
     errs = pd.Series(errs)
 
     stat = {
-        'sd_duration' : df['duration'].std(),
+        'sd_duration' : df['duration'].std() / 60,
         'mean_err': errs.mean(),
         'sd_err': errs.std(),
         'mean_abs_err': errs.abs().mean(),

### 2017-12-09 reports/paper/paper.tex
.

@@
 Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields the following lower bound for negative log-likelihood, up to a constant,
-\begin{align*}
-&\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} + \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i \\&\qquad + \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i},
-\end{align*}
-where we sample uniformly and independently $\{\tilde Z_j\uppr{k}\}_{j=1}^L$.
+\begin{align}
+&\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} + \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i 
+\label{eq:softmax_objective}
+\\&\qquad + \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i} \nonumber,
+\end{align}
+where we sample uniformly and independently $\{\tilde Z_j\uppr{k}\}_{j=1}^L$. The pseudocode of the implementation is detailed in Algorithm~\ref{alg:main_alg}.
 
@@
   \begin{algorithmic}
-    \STATE{}
+    \STATE{\COMMENT{Optimizing using Adam in PyTorch}}
+
+    \FOR{trip $k$ in $1,\ldots,N$} \STATE {
+    Sample $L$ random paths uniformly.
+
+    Compute the objective function as in~\eqref{eq:softmax_objective} for observation $k$.
+
+    Compute the gradients of the objective function with respect to $\bm W$.
+
+    Update $\bm W$.
+    } \ENDFOR
   \end{algorithmic}
-  \caption{Your Pseudocode}
+  \caption{Training algorithm for softmax path selection}
+  \label{alg:main_alg}
 \end{algorithm}

### 2017-12-09 test.py
Properly converged ncr unif model.

@@ -21,7 +21,7 @@ def test(predict, result_file,
     errs = pd.Series(errs)
 
     stat = {
-        'sd_duration' : df['duration'].std() / 60,
+        'sd_duration' : df['duration'].std(),
         'mean_err': errs.mean(),
         'sd_err': errs.std(),
         'mean_abs_err': errs.abs().mean(),

### 2017-12-09 reports/paper/paper.tex
Properly converged ncr unif model.
.

@@
 \usepackage{algorithm}
-\usepackage{algorithmic}
+\usepackage{algorithmic,inconsolata}
 \usepackage{lipsum}
@@
 \usepackage{natbib}
+\usepackage{threeparttable}
+\newcommand{\sd}{\operatorname{SD}}
 
@@
 
-{\color{gray}
-\begin{itemize}
-\item How do you plan on training your parameters / inferring the
-  states of your latent variables (MLE / MAP / Backprop / VI / EM / BP / ...)
-
-\item What are the assumptions implicit in this technique? Is it an approximation or exact? If it is an approximation what bound does it optimize?
-
-\item What is the explicit method / algorithm that you derive for learning these parameters?
-\end{itemize}
-
 \begin{algorithm}
   \begin{algorithmic}
-    \STATE{\COMMENT{Optimizing using Adam in PyTorch}}
+    \STATE{\COMMENT{Optimizing using \texttt{torch.optim.Adam}}}
 
@@
 \end{algorithm}
-}
+
+\subsection{Prediction}
+\label{sub:predict}
+We conclude this section by briefly discussing prediction in our models.
+In the uniform model, prediction is straightforward.
+We calculate the expected time $\hat T = \sum_i w_i \pi_i$, where $\pi_i$ is the marginal probability that node $i$ is included in a uniformly randomly chosen path.
+Computing prediction is slightly more subtle in the softmax regression model.
+
+% {\color{gray}
+% \begin{itemize}
+% \item How do you plan on training your parameters / inferring the
+%   states of your latent variables (MLE / MAP / Backprop / VI / EM / BP / ...)
+
+% \item What are the assumptions implicit in this technique? Is it an approximation or exact? If it is an approximation what bound does it optimize?
+
+% \item What is the explicit method / algorithm that you derive for learning these parameters?
+% \end{itemize}
+
+% }
 
@@
 
-\begin{itemize}
+{\color{gray}\begin{itemize}
 \item What were the results comparing previous work / baseline systems / your systems on the main task?
@@
 \end{itemize}
+}
+
+We train our model and
+
+Table~\ref{tab:mainres} contains the main prediction results of our model relative to some benchmark predictions
+such as the Google Maps Distance Matrix API, neural network, and linear regression.
 
@@
 \centering
-  \missingfigure{}
-  \caption{This is usually a table. Tables with numbers are generally easier to read than graphs, so prefer when possible.}
-  \label{fig:mainres}
+\begin{threeparttable}
+\begin{tabular}{lrrrrr}
+\toprule
+& \multicolumn{3}{c}{Baseline} & \multicolumn{2}{c}{Model} \\\cmidrule(r){2-4} \cmidrule(l){5-6} 
+{} & Google Maps & Neural Network & Linear Regression & Uniform Model & Softmax Model \\
+\midrule
+$\sd(T)$ & 7.40 & 6.85& &  6.22 & 6.22 \\ [1em]
+$\hat \E[\epsilon]$ & -0.83 &-0.07 & &  0.61 & 0.99 \\
+$\sd(\epsilon)$ & 4.98 & 3.88& &  4.28 & 4.28 \\[1em]
+$\hat \E[|\epsilon|]$ & 3.42 & 2.44& &  2.82 & 2.86 \\
+$\text{Median}(|\epsilon|)$ & 2.55 & 1.67& &  1.92 & 1.92 \\
+$\text{Percentile}(|\epsilon|, 99)$ & 16.54 &12.79 & &  14.72 & 15.24 \\ [1em]
+Test-set $R^2$ & 0.54 & 0.68&  & 0.53 & 0.53 \\
+\bottomrule
+\end{tabular}
+\caption{Prediction results on test set}
+\label{tab:mainres}
+\begin{tablenotes}
+\item \textbf{Notes:} Here $\epsilon = (\text{Actual trip duration}) - (\text{Predicted duration})$ in minutes.
+$\hat \E$ denotes the sample mean.
+We restrict the test-set to 8:00--9:00 AM on Tuesdays in January 2009 for all models \emph{except} the neural network, for which we use an out-of-sample test set on the \emph{entire} range of trip times, since the neural network encodes time-of-day and day-of-week data.
+Test-set $R^2$ is calculated as \smash{$1-\frac{\var(\epsilon)}{\var(T)}$} on the underlying dataset being tested on.
+Note that for all but the linear regression, the test-set $R^2$ is an out-of-sample $R^2$, whereas the linear regression is trained on the test set, and thus the $R^2$ reported is in-sample.
+\end{tablenotes}
+\end{threeparttable}
 \end{table*}

### 2017-12-09 reports/paper/paper.tex
results_1

@@
 Computing prediction is slightly more subtle in the softmax regression model.
+In the softmax regression model, we sample $L_{\text{pred}}$ uniformly chosen paths and take the weighted average of the sums of $w_i$ in each path, where the weights are proportional to $\exp(-\sum_{i \in Z} w_i)$ for path $Z$.
 
@@
 
-We train our model and
+We train our model and test the model on an out-of-sample test-set using the methods discussed in Section~\ref{sub:predict}.
 
@@
 & \multicolumn{3}{c}{Baseline} & \multicolumn{2}{c}{Model} \\\cmidrule(r){2-4} \cmidrule(l){5-6} 
-{} & Google Maps & Neural Network & Linear Regression & Uniform Model & Softmax Model \\
+{} & Google Maps & Neural Network\tnote{2} & Linear Regression\tnote{3} & Uniform Model & Softmax Model \\
 \midrule
-$\sd(T)$ & 7.40 & 6.85& &  6.22 & 6.22 \\ [1em]
-$\hat \E[\epsilon]$ & -0.83 &-0.07 & &  0.61 & 0.99 \\
-$\sd(\epsilon)$ & 4.98 & 3.88& &  4.28 & 4.28 \\[1em]
-$\hat \E[|\epsilon|]$ & 3.42 & 2.44& &  2.82 & 2.86 \\
-$\text{Median}(|\epsilon|)$ & 2.55 & 1.67& &  1.92 & 1.92 \\
-$\text{Percentile}(|\epsilon|, 99)$ & 16.54 &12.79 & &  14.72 & 15.24 \\ [1em]
-Test-set $R^2$ & 0.54 & 0.68&  & 0.53 & 0.53 \\
+$\sd(T)$ & 7.40 & 6.85& 6.13&  6.22 & 6.22 \\ [1em]
+$\hat \E[\epsilon]\tnote{1}$ & -0.83 &-0.07 & -0.05&  0.61 & 0.99 \\
+$\sd(\epsilon)$ & 4.98 & 3.88& 5.01&  4.28 & 4.28 \\[1em]
+$\hat \E[|\epsilon|]$ & 3.42 & 2.44& 3.60&  2.82 & 2.86 \\
+$\text{Median}(|\epsilon|)$ & 2.55 & 1.67& 2.93 &  1.92 & 1.92 \\
+$\text{Percentile}(|\epsilon|, 99)$ & 16.54 &12.79 & 16.62&  14.72 & 15.24 \\ [1em]
+Test-set\tnote{4}  \, $R^2$ & 0.54 & 0.68& 0.33 & 0.53 & 0.53 \\
 \bottomrule
@@
 \begin{tablenotes}
-\item \textbf{Notes:} Here $\epsilon = (\text{Actual trip duration}) - (\text{Predicted duration})$ in minutes.
-$\hat \E$ denotes the sample mean.
-We restrict the test-set to 8:00--9:00 AM on Tuesdays in January 2009 for all models \emph{except} the neural network, for which we use an out-of-sample test set on the \emph{entire} range of trip times, since the neural network encodes time-of-day and day-of-week data.
-Test-set $R^2$ is calculated as \smash{$1-\frac{\var(\epsilon)}{\var(T)}$} on the underlying dataset being tested on.
-Note that for all but the linear regression, the test-set $R^2$ is an out-of-sample $R^2$, whereas the linear regression is trained on the test set, and thus the $R^2$ reported is in-sample.
+% \item[\textbf{Notes:}]
+\item[1] Here $\epsilon = (\text{Actual trip duration}) - (\text{Predicted duration})$ in minutes. $\hat \E$ denotes the sample mean. 
+\item[2] We restrict the test-set to 8:00--9:00 AM on Tuesdays in January 2009 for all models \emph{except} the neural network, for which we use an out-of-sample test set on the \emph{entire} range of trip times, since the neural network encodes time-of-day and day-of-week data.
+\item[3]In linear regression, we simply regress $T$ on straight-line distance between the source and the destination.  
+\item[4] Test-set $R^2$ is calculated as \smash{$1-\frac{\var(\epsilon)}{\var(T)}$} on the underlying dataset being tested on. Note that the test-set $R^2$ is an out-of-sample $R^2$. 
 \end{tablenotes}

### 2017-12-09 reports/paper/paper.tex
Add notebooks and paper background.

--- 2017-12-09 reports/paper/paper.tex
@@
 
+}
 \begin{algorithm}
   \begin{algorithmic}
-    \STATE{\COMMENT{Optimizing using Adam in PyTorch}}
+    \STATE{\COMMENT{Optimizing using \texttt{torch.optim.Adam}}}
 
@@
 \end{algorithm}
-}
 
@@
 
+\subsection{Data Processing}
+
 For our dataset, we started with the TLC dataset of all NYC taxi trips taken in
@@
 
-In addition to adding columns, the original dataset was filtered. Geographic
-coordinates were\ldots
+In addition to adding columns, the original dataset was filtered. Only trips
+whose start and locations fell inside a rectangle that bounds Manhattan were
+considered. As a preliminary filtering, trips with duration less than 0 or
+greater than one day were removed. Upon training models, this was further
+culled to trips between two minutes and two hours in length.
+
+Principal component analysis was run on all de-meaned and
+standardized geographic coordinates (both start
+and end locations) from the filtered data. The coordinates were thus
+transformed into the first and second principal components. For the discrete
+grid models, these coordinates were further discretized into a $20 \times 70$
+grid.
+
+The data-set was saved as a whole, but also partioned into $(\text{day of week},
+\text{hour})$ groups, using the one-hot encoded vectors. Some models were
+trained on the entire data-set (e.g.\ Neural Network), and some were trained
+individually on the partioned data-sets (e.g.\ Structural Models).
+
+\subsection{Baselines}
+
+Three baselines were calculated. The first was a linear regression predicting
+trip duration from trip distance (calculated using the Pythagorean Theorem and
+the start and end coordinates), and trained on a partitioned data set.
+
+The second was a neural network trained on the
+entire dataset taking as input the start and end coordinates as well as the day
+of week and hour one-hot encoded vectors. The neural network had one hidden
+layer with 50 neurons and a tanh activation function.
+
+The second baseline was constructed by querying the Google Maps Distance Matrix
+API\footnote{\url{https://developers.google.com/maps/documentation/distance-matrix/}}.
+While there was no need to train this baseline, there was a rate limit
+constraint on how many test data points it could be tested on. We assess its
+performance on 2,500 data points in a representative partioned test data-set.
+
+\subsection{Training}
+
+Training implementations varied from model to model. In training the neural
+network baseline, SGD was used with a learning rate of $10^{-7}$ run for 12
+epochs on the full data set. This took between 30 minutes and an hour. The
+uniform paths structural model was similarly trained with SGD, with a learning
+rate of $10^{-4}$.
+
+The uniform model was only trained on a subset of the
+dataset corresponding to all trips on a given day of the week and on a given
+hour. Convergence appeared after roughly two epochs, and the model was run for 5
+epochs taking just under ten minutes.
+
+The softmax structural model was trained
+using the \texttt{Adam} optimizer with learning rate $10^{-2}$ for 10 epochs,
+taking approximately an hour.
+% TODO: jchen confirm this for me:
 

### 2017-12-09 reports/paper/paper.tex
edited methods

--- 2017-12-09 reports/paper/paper.tex
@@
 
+\subsection{Data Processing}
+
 For our dataset, we started with the TLC dataset of all NYC taxi trips taken in
-January 2009. From the raw data, several computed columns were added.
-One-hot-encoded columns were added for hour of day (24 columns) as well as for
-day of the week (7 columns). The duration of the trip was also explicitly
-computed in seconds as the dependent variable to predict.
+January 2009. From the raw data, we added several computed columns.
+We one-hot-encoded columns for hour of day (24 columns) as well as for
+day of the week (7 columns). We also explicitly computed the duration of the trip  in seconds as the dependent variable to predict.
 
-In addition to adding columns, the original dataset was filtered. Geographic
-coordinates were\ldots
+In addition to adding columns, we filtered the original dataset. We only consider trips
+whose start- and end-locations that falls inside a rectangle that bounds Manhattan. We also removed trips with duration less than two minutes or more than two hours.
+
+We performed a principal component analysis on all de-meaned and
+standardized geographic coordinates (both start
+and end locations) from the filtered data.
+We thus transform the coordinates into the first and second principal components.
+This allowed us to compute the rotational matrix so that Manhattan lies approximately orthogonal to the $x$-axis.
+For the discrete
+grid models, these coordinates were further discretized into a $20 \times 70$
+grid. Each node in the grid approximately corresponds to a $400\text{ft} \times 400\text{ft}$ area. We chose the grid size as a compromise between computational tractability and model expressiveness.
+
+We saved the dataset as a whole, and we also partitioned into $(\text{day of week},
+\text{hour})$ groups, using the one-hot encoded vectors. Some models were
+trained on the entire dataset (e.g.\ Neural Network), and some were trained
+individually on the partioned datasets (e.g.\ Structural Models).
+
+\subsection{Baselines}
+
+We calculated three baselines. The first was a linear regression predicting
+trip duration from trip distance (calculated using the Pythagorean Theorem and
+the start and end coordinates), and trained on a partitioned data set.
+
+The second was a neural network trained on the
+entire dataset taking as input the start and end coordinates as well as the day
+of week and hour one-hot encoded vectors. The neural network had one hidden
+layer with 50 neurons and a $\tanh$ activation function.
+
+The third baseline was constructed by querying the Google Maps Distance Matrix
+API.\footnote{\url{https://developers.google.com/maps/documentation/distance-matrix/}}
+While there was no need to train this baseline, there was a rate limit
+constraint on how many test data points it could be tested on. We assess its
+performance on 2,500 data points in a representative partioned test dataset.
+
+\subsection{Training}
+
+Training implementations varied from model to model. In training the neural
+network baseline, SGD was used with a learning rate of $10^{-7}$ run for 12
+epochs on the full data set. This took between 30 minutes and an hour. The
+uniform paths structural model was similarly trained with SGD, with a learning
+rate of $10^{-4}$.
+
+The uniform model was only trained on a subset of the
+dataset corresponding to all trips on a given day of the week and on a given
+hour. Convergence appeared after roughly two epochs, and the model was run for 5
+epochs taking just under ten minutes.
+
+The softmax regression model was trained
+using the Adam optimizer \cite{kingma2014adam} with learning rate $10^{-2}$ for 10 epochs,
+taking approximately an hour.
+Additionally, the softmax regression model take hyperparameters $L$ (sample size in each stochastic draw) and $\sigma^2$ (variance of trip time).
+We chose $L = 20$ and $\sigma^2 = 100$.
+These values are again chosen to strike a bargain between realism and computation feasibility.
+% TODO: jchen confirm this for me:
 
@@
 
-\begin{table*}
+\begin{table*}[h!]
 \centering

### 2017-12-09 reports/paper/paper.tex
.

@@
 This is the same technique as in importance sampling, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations.
-Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields the following lower bound for negative log-likelihood, up to a constant,
+Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields the following upper bound for negative log-likelihood, up to a constant,
 \begin{align}
@@
 These values are again chosen to strike a bargain between realism and computation feasibility.
-% TODO: jchen confirm this for me:
+The model appears to converge by the second epoch.
 
@@
 We train our model and test the model on an out-of-sample test-set using the methods discussed in Section~\ref{sub:predict}.
-
 Table~\ref{tab:mainres} contains the main prediction results of our model relative to some benchmark predictions
 such as the Google Maps Distance Matrix API, neural network, and linear regression.
+We note that, in terms of predictive metrics, our models were able to outperform Google Maps and linear regression, but were not able to outperform the neural network, even though the difference is not very large.
+Despite the disappointing outcome, the advantage of our framework is that the parameters have structural interpretations, whereas the neural network is a black box.
+
+In terms of prediction, the uniform model and the softmax model appear very similar, whereas the uniform model performs slightly better.
+We suspect that, due to computational constraints, $L$ is too small for the softmax model to really be expressive, as the number of weights updated by the softmax model is much fewer than that of the uniform model.
+Curiously, both the uniform and softmax models have a nontrivial bias in $\hat\E(\epsilon)$.
+We suspect that the bias is generated by the approximate nature of our inference procedures, where our objective functions are approximations to the likelihood.
+
+% TODO fig
 

### 2017-12-10 reports/paper/paper.tex
Add learned weight data, put unif map figure into paper.
Added softmax loss.
.
Clean up paper.

@@
 
-{\color{gray}\begin{itemize}
-\item What were the results comparing previous work / baseline systems / your systems on the main task?
-\item What were the secondary results comparing the variants of your system?
-\item This section should be fact based and relatively dry. What happened, what was significant?
-\end{itemize}
-}
+% {\color{gray}\begin{itemize}
+% \item What were the results comparing previous work / baseline systems / your systems on the main task?
+% \item What were the secondary results comparing the variants of your system?
+% \item This section should be fact based and relatively dry. What happened, what was significant?
+% \end{itemize}
+% }
 
@@
 We suspect that the bias is generated by the approximate nature of our inference procedures, where our objective functions are approximations to the likelihood.
-
-% TODO fig
 
@@
 
-% \begin{table}
-%   \centering
-%   \missingfigure[figheight=5cm]{}
-%   \caption{Secondary table or figure in results section.}
-%   \label{fig:mainres}
-% \end{table}
-
 \section{Discussion}
 
-\begin{itemize}
-\item What conclusions can you draw from the results section?
-\item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
+% \begin{itemize}
+% \item What conclusions can you draw from the results section?
+% \item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
+% \item  What questions remain open? What did you think might work, but did not?
+% \end{itemize}
 
-\item  What questions remain open? What did you think might work, but did not?
-\end{itemize}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{figs/unifmaps.pdf}
+\includegraphics[width=0.5\textwidth]{figs/softmaxmaps.pdf}
+\caption{Learned Weights, Various Times Tuesday. Top Row: Uniform Model.
+Bottom Row: Softmax Model.}
+\label{fig:unifmap}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{figs/softmaxloss}
+\caption{Softmax model training loss, rolling average.}
+\label{fig:softmaxloss}
+\end{figure}
 

### 2017-12-10 reports/paper/paper.tex
related work

--- 2017-12-09 reports/paper/paper.tex
@@
 \usepackage{algorithm}
-\usepackage{algorithmic,inconsolata}
+\usepackage{algorithmic}
 \usepackage{lipsum}
@@
 \usepackage{natbib}
-\usepackage{threeparttable}
-\newcommand{\sd}{\operatorname{SD}}
 
@@
 \end{itemize}
+
+Travel time prediction has historically been a topic of research interest.
+On the topic of freeway travel time prediction, \cite{van2002freeway} proposed a Recurrent Neural Network (RNN) approach in order to address the temporal aspect of traffic prediction; \cite{wu2004travel} adopted the Supportive Vector Regression (SVR) method applied to time-seires analysis and reduced prediction errors in cases where previous methods generate especially large errors.
+We notice that earlier papers share the common feature of using freeway data for training and prediction, which probably arised from the limited availability of GPS data sources.
+Nevertheless, we realize that even for freeways, which in general have less traffic conditions than city roads, there is evidence that there exists a high level of non-linearity in the data, which contributes to the good performance of models such as RNN and SVR.
+
+With the ubiquity of taxi GPS data, recent work focused on estimating the traffic conditions in cities using sparce probe data and used travel time predictions as the metric as the performance of the model.
+\cite{hunter2009path} uses a Bayesian framework and used an expectation maximization algorithm that simultaneously learns the likely paths taken by probe vehicles as well as the travel time distributions through the network; \cite{herring2010estimating} used a Coupled Hidden Markov Model (CHMM) and determines the most likely path by allocating the travel time between two consecutive location observations to roads in the `M' step.
+
+One thing that distinguishes our attempt to the previous work is that we have much more missing information.
+Instead of sparse probe data, we are only working with start and end locations, which presents much more challenges in actual path inference and parameter optimization.
+\cite{zhan2013urban} provided some very helpful insights since it used the same TLC dataset of taxi ride.
+It contructed a faithful representation of the manhattan streets network, but limited the discussion to a much smaller district in Manhattan to reduce computational complexity.
 
@@
 
+{\color{gray}
+\begin{itemize}
+\item How do you plan on training your parameters / inferring the
+  states of your latent variables (MLE / MAP / Backprop / VI / EM / BP / ...)
+
+\item What are the assumptions implicit in this technique? Is it an approximation or exact? If it is an approximation what bound does it optimize?
+
+\item What is the explicit method / algorithm that you derive for learning these parameters?
+\end{itemize}
+
 \begin{algorithm}
   \begin{algorithmic}
-    \STATE{\COMMENT{Optimizing using \texttt{torch.optim.Adam}}}
+    \STATE{\COMMENT{Optimizing using Adam in PyTorch}}
 
@@
 \end{algorithm}
-
-\subsection{Prediction}
-\label{sub:predict}
-We conclude this section by briefly discussing prediction in our models.
-In the uniform model, prediction is straightforward.
-We calculate the expected time $\hat T = \sum_i w_i \pi_i$, where $\pi_i$ is the marginal probability that node $i$ is included in a uniformly randomly chosen path.
-Computing prediction is slightly more subtle in the softmax regression model.
-In the softmax regression model, we sample $L_{\text{pred}}$ uniformly chosen paths and take the weighted average of the sums of $w_i$ in each path, where the weights are proportional to $\exp(-\sum_{i \in Z} w_i)$ for path $Z$.
-
-% {\color{gray}
-% \begin{itemize}
-% \item How do you plan on training your parameters / inferring the
-%   states of your latent variables (MLE / MAP / Backprop / VI / EM / BP / ...)
-
-% \item What are the assumptions implicit in this technique? Is it an approximation or exact? If it is an approximation what bound does it optimize?
-
-% \item What is the explicit method / algorithm that you derive for learning these parameters?
-% \end{itemize}
-
-% }
+}
 
@@
 
-\subsection{Data Processing}
+For our dataset, we started with the TLC dataset of all NYC taxi trips taken in
+January 2009. From the raw data, several computed columns were added.
+One-hot-encoded columns were added for hour of day (24 columns) as well as for
+day of the week (7 columns). The duration of the trip was also explicitly
+computed in seconds as the dependent variable to predict.
 
-For our dataset, we started with the TLC dataset of all NYC taxi trips taken in
-January 2009. From the raw data, we added several computed columns.
-We one-hot-encoded columns for hour of day (24 columns) as well as for
-day of the week (7 columns). We also explicitly computed the duration of the trip  in seconds as the dependent variable to predict.
-
-In addition to adding columns, we filtered the original dataset. We only consider trips
-whose start- and end-locations that falls inside a rectangle that bounds Manhattan. We also removed trips with duration less than two minutes or more than two hours.
-
-We performed a principal component analysis on all de-meaned and
-standardized geographic coordinates (both start
-and end locations) from the filtered data.
-We thus transform the coordinates into the first and second principal components.
-This allowed us to compute the rotational matrix so that Manhattan lies approximately orthogonal to the $x$-axis.
-For the discrete
-grid models, these coordinates were further discretized into a $20 \times 70$
-grid. Each node in the grid approximately corresponds to a $400\text{ft} \times 400\text{ft}$ area. We chose the grid size as a compromise between computational tractability and model expressiveness.
-
-We saved the dataset as a whole, and we also partitioned into $(\text{day of week},
-\text{hour})$ groups, using the one-hot encoded vectors. Some models were
-trained on the entire dataset (e.g.\ Neural Network), and some were trained
-individually on the partioned datasets (e.g.\ Structural Models).
-
-\subsection{Baselines}
-
-We calculated three baselines. The first was a linear regression predicting
-trip duration from trip distance (calculated using the Pythagorean Theorem and
-the start and end coordinates), and trained on a partitioned data set.
-
-The second was a neural network trained on the
-entire dataset taking as input the start and end coordinates as well as the day
-of week and hour one-hot encoded vectors. The neural network had one hidden
-layer with 50 neurons and a $\tanh$ activation function.
-
-The third baseline was constructed by querying the Google Maps Distance Matrix
-API.\footnote{\url{https://developers.google.com/maps/documentation/distance-matrix/}}
-While there was no need to train this baseline, there was a rate limit
-constraint on how many test data points it could be tested on. We assess its
-performance on 2,500 data points in a representative partioned test dataset.
-
-\subsection{Training}
-
-Training implementations varied from model to model. In training the neural
-network baseline, SGD was used with a learning rate of $10^{-7}$ run for 12
-epochs on the full data set. This took between 30 minutes and an hour. The
-uniform paths structural model was similarly trained with SGD, with a learning
-rate of $10^{-4}$.
-
-The uniform model was only trained on a subset of the
-dataset corresponding to all trips on a given day of the week and on a given
-hour. Convergence appeared after roughly two epochs, and the model was run for 5
-epochs taking just under ten minutes.
-
-The softmax regression model was trained
-using the Adam optimizer \cite{kingma2014adam} with learning rate $10^{-2}$ for 10 epochs,
-taking approximately an hour.
-Additionally, the softmax regression model take hyperparameters $L$ (sample size in each stochastic draw) and $\sigma^2$ (variance of trip time).
-We chose $L = 20$ and $\sigma^2 = 100$.
-These values are again chosen to strike a bargain between realism and computation feasibility.
-% TODO: jchen confirm this for me:
+In addition to adding columns, the original dataset was filtered. Geographic
+coordinates were\ldots
 
@@
 
-{\color{gray}\begin{itemize}
+\begin{itemize}
 \item What were the results comparing previous work / baseline systems / your systems on the main task?
@@
 \end{itemize}
-}
 
-We train our model and test the model on an out-of-sample test-set using the methods discussed in Section~\ref{sub:predict}.
-
-Table~\ref{tab:mainres} contains the main prediction results of our model relative to some benchmark predictions
-such as the Google Maps Distance Matrix API, neural network, and linear regression.
-
-\begin{table*}[h!]
+\begin{table*}
   \centering
-\begin{threeparttable}
-\begin{tabular}{lrrrrr}
-\toprule
-& \multicolumn{3}{c}{Baseline} & \multicolumn{2}{c}{Model} \\\cmidrule(r){2-4} \cmidrule(l){5-6} 
-{} & Google Maps & Neural Network\tnote{2} & Linear Regression\tnote{3} & Uniform Model & Softmax Model \\
-\midrule
-$\sd(T)$ & 7.40 & 6.85& 6.13&  6.22 & 6.22 \\ [1em]
-$\hat \E[\epsilon]\tnote{1}$ & -0.83 &-0.07 & -0.05&  0.61 & 0.99 \\
-$\sd(\epsilon)$ & 4.98 & 3.88& 5.01&  4.28 & 4.28 \\[1em]
-$\hat \E[|\epsilon|]$ & 3.42 & 2.44& 3.60&  2.82 & 2.86 \\
-$\text{Median}(|\epsilon|)$ & 2.55 & 1.67& 2.93 &  1.92 & 1.92 \\
-$\text{Percentile}(|\epsilon|, 99)$ & 16.54 &12.79 & 16.62&  14.72 & 15.24 \\ [1em]
-Test-set\tnote{4}  \, $R^2$ & 0.54 & 0.68& 0.33 & 0.53 & 0.53 \\
-\bottomrule
-\end{tabular}
-\caption{Prediction results on test set}
-\label{tab:mainres}
-\begin{tablenotes}
-% \item[\textbf{Notes:}]
-\item[1] Here $\epsilon = (\text{Actual trip duration}) - (\text{Predicted duration})$ in minutes. $\hat \E$ denotes the sample mean. 
-\item[2] We restrict the test-set to 8:00--9:00 AM on Tuesdays in January 2009 for all models \emph{except} the neural network, for which we use an out-of-sample test set on the \emph{entire} range of trip times, since the neural network encodes time-of-day and day-of-week data.
-\item[3]In linear regression, we simply regress $T$ on straight-line distance between the source and the destination.  
-\item[4] Test-set $R^2$ is calculated as \smash{$1-\frac{\var(\epsilon)}{\var(T)}$} on the underlying dataset being tested on. Note that the test-set $R^2$ is an out-of-sample $R^2$. 
-\end{tablenotes}
-\end{threeparttable}
+  \missingfigure{}
+  \caption{This is usually a table. Tables with numbers are generally easier to read than graphs, so prefer when possible.}
+  \label{fig:mainres}
 \end{table*}

### 2017-12-10 reports/paper/paper.tex
.

--- 2017-12-10 reports/paper/paper.tex
@@
 
-\section{Inference (or Training)}
+\section{Inference}
 \label{sec:inference}
@@
 
-\subsection{Inference in the uniform model}
+\subsection{Uniform Model}
 
@@
 
-\subsection{Inference in the softmax regression model}
+\subsection{Softmax Regression Model}
 
@@
 
-% {\color{gray}\begin{itemize}
-% \item What were the results comparing previous work / baseline systems / your systems on the main task?
-% \item What were the secondary results comparing the variants of your system?
-% \item This section should be fact based and relatively dry. What happened, what was significant?
-% \end{itemize}
-% }
+{\color{gray}\begin{itemize}
+\item What were the results comparing previous work / baseline systems / your systems on the main task?
+\item What were the secondary results comparing the variants of your system?
+\item This section should be fact based and relatively dry. What happened, what was significant?
+\end{itemize}
+}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{figs/softmaxloss}
+\caption{Softmax model training loss, rolling average.}
+\label{fig:softmaxloss}
+\end{figure}
 
 We train our model and test the model on an out-of-sample test-set using the methods discussed in Section~\ref{sub:predict}.
+As a simple check of the training process, we plot in Figure~\ref{fig:softmaxloss} the loss obtained when fitting the softmax regression model.
+We see that the model converges relatively fast.\footnote{The full fitting takes approximately 30 minutes per dataset (2.9 GHz Intel Core i5, 8GB RAM), but the model appears to converge in only 20\% of that time.} The convergence plot for the uniform model is extremely similar and thus omitted.
 Table~\ref{tab:mainres} contains the main prediction results of our model relative to some benchmark predictions
@@
 We suspect that, due to computational constraints, $L$ is too small for the softmax model to really be expressive, as the number of weights updated by the softmax model is much fewer than that of the uniform model.
-Curiously, both the uniform and softmax models have a nontrivial bias in $\hat\E(\epsilon)$.
-We suspect that the bias is generated by the approximate nature of our inference procedures, where our objective functions are approximations to the likelihood.
+% TODO fig
 
@@
 
+% \begin{table}
+%   \centering
+%   \missingfigure[figheight=5cm]{}
+%   \caption{Secondary table or figure in results section.}
+%   \label{fig:mainres}
+% \end{table}
+
 \section{Discussion}
 
-% \begin{itemize}
-% \item What conclusions can you draw from the results section?
-% \item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
-% \item  What questions remain open? What did you think might work, but did not?
-% \end{itemize}
+% table
+%
 
-\begin{figure}[h]
+Table~\ref{tab:mainres} is encouraging.
+Our models successfully leverage the power of geography and picks up more signal in the data than distance of the trip alone, as demonstrated by the significantly improved performance over linear regression.
+Perhaps more strikingly, our model slightly outperforms the predictions generated by Google Maps, which means that the predictive accuracy of the model is on par with the standard of a widely used consumer product.
+Even though we were not able to outperform the neural network, we reiterate that the main advantage of our approach is to have interpretable parameters, which are visualized in Figure~\ref{fig:unifmap}.
+
+\begin{figure}[h!]
 \centering
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/softmaxloss}
-\caption{Softmax model training loss, rolling average.}
-\label{fig:softmaxloss}
-\end{figure}
+\subsection{Comparing Models}
+
+\subsection{Remaining Puzzles}
+
+\begin{itemize}
+\item What conclusions can you draw from the results section?
+\item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
+
+\item  What questions remain open? What did you think might work, but did not?
+\end{itemize}
 

### 2017-12-10 reports/paper/paper.tex

--- 2017-12-10 reports/paper/paper.tex
@@
 Despite the disappointing outcome, the advantage of our framework is that the parameters have structural interpretations, whereas the neural network is a black box.
-
-In terms of prediction, the uniform model and the softmax model appear very similar, whereas the uniform model performs slightly better.
-We suspect that, due to computational constraints, $L$ is too small for the softmax model to really be expressive, as the number of weights updated by the softmax model is much fewer than that of the uniform model.
-Curiously, both the uniform and softmax models have a nontrivial bias in $\hat\E(\epsilon)$.
-We suspect that the bias is generated by the approximate nature of our inference procedures, where our objective functions are approximations to the likelihood.
 
@@
 
+% table
+%
+
+Table~\ref{tab:mainres} is encouraging.
+Our models successfully leverage the power of geography and picks up more signal in the data than distance of the trip alone, as demonstrated by the significantly improved performance over linear regression.
+Perhaps more strikingly, our model slightly outperforms the predictions generated by Google Maps, which means that the predictive accuracy of the model is on par with the standard of a widely used consumer product.
+Even though we were not able to outperform the neural network, we reiterate that the main advantage of our approach is to have interpretable parameters, which are visualized in Figure~\ref{fig:unifmap}.
+=======
 % \begin{itemize}
@@
 % \end{itemize}
+>>>>>>> 47a18b762f9ece54e1695ac601a050378593adc7
 
-\begin{figure}[h]
+\begin{figure}[h!]
 \centering
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/softmaxloss}
-\caption{Softmax model training loss, rolling average.}
-\label{fig:softmaxloss}
-\end{figure}
+\subsection{Comparing Models}
+
+\subsection{Remaining Puzzles}
+
+\begin{itemize}
+\item What conclusions can you draw from the results section?
+\item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
+
+\end{itemize}
 

### 2017-12-10 reports/paper/paper.tex
merged conflict resolved
.

@@
 \begin{abstract}
-  \begin{itemize}
-  \item This document describes the expected style, structure, and rough proportions for your final project write-up.
-  \item While you are free to break from this structure, consider it a strong prior for our expectations of the final report.
-\item Length is a hard constraint.
+%   \begin{itemize}
+%   \item This document describes the expected style, structure, and rough proportions for your final project write-up.
+%   \item While you are free to break from this structure, consider it a strong prior for our expectations of the final report.
+% \item Length is a hard constraint.
 You are only allowed max \textbf{8 pages} in this format.
@@
 It is your responsibility to convey the main contributions of the work in the length given.
-  \end{itemize}
+%   \end{itemize}
 
@@
 
-Example Structure:
-\begin{itemize}
-\item What is the problem of interest and what (high-level) are the current best methods for solving it?
-\item How do you plan to improve/understand/modify this or related methods?
-\item Preview your research process, list the contributions you made, and summarize your experimental findings.
-\end{itemize}
+% re
 
@@
 \includegraphics[width=0.5\textwidth]{figs/ridesbyhour.pdf}
-\caption{Taxi Trips by Time of Day}
+\caption{Taxi trips by time of day}
 \label{fig:ridesbyhour}
@@
 \includegraphics[width=0.5\textwidth]{figs/tripdurations.pdf}
-\caption{Taxi Trips by Duration}
+\caption{Taxi trips by duration}
 \label{fig:tripdurations}
@@
 
-Example Structure:
-\begin{itemize}
-\item What 3-5 papers have been published in this space?
-\item How do these  differ from your approach?
-\item What data or methodologies do each of these works use?
-\item How do you plan to compare to these methods?
-\end{itemize}
+% Example Structure:
+% \begin{itemize}
+% \item What 3-5 papers have been published in this space?
+% \item How do these  differ from your approach?
+% \item What data or methodologies do each of these works use?
+% \item How do you plan to compare to these methods?
+% \end{itemize}
 
@@
 \includegraphics[width=0.5\textwidth]{figs/softmaxloss}
-\caption{Softmax model training loss, rolling average.}
+\caption{Softmax model training loss, rolling average}
 \label{fig:softmaxloss}
@@
 Even though we were not able to outperform the neural network, we reiterate that the main advantage of our approach is to have interpretable parameters, which are visualized in Figure~\ref{fig:unifmap}.
-=======
+We turn to the figure in the next section.
 % \begin{itemize}
@@
 % \end{itemize}
->>>>>>> 47a18b762f9ece54e1695ac601a050378593adc7
 
@@
 
+Figure~\ref{fig:unifmap} shows that the parameter values for the uniform and softmax regression models behave similarly, with similar areas of high and low congestion.
+The patterns seem to largely agree with conventional wisdom, both in spatial and temporal dimensions.
+We identify that 4:00 AM is indeed less congested than rush hour, that Midtown Manhattan seems to be particularly congested during the day, and that the congested areas have an average speed of 4--6 miles per hour.
+The visualization also confirms some less-than-well-known features; the areas that demonstrate congestion at 4:00 AM largely confirms with anecdotal evidence.\footnote{According to \textit{Business Insider} (\url{http://read.bi/164vSK1}), East Village and surrounding areas rank highly for nightlife.
+%#GETLIT
+} As the models seem to produce largely similar output, we conclude that they are fairly robust.
+
+However, the two models do have significant differences.
+The softmax regression model appears to have smoother parameter values than the uniform model.
+This is consistent with the softmax model ascribing higher probability to faster paths, and so we expect the parameter updates to spread out spatially.
+
+It is difficult to select the model to use based on predictive metrics alone, as the models perform strikingly similarly.
+However, from a practical standpoint, the uniform model is significantly easier to implement and faster to train than the softmax regression model, even though the latter embraces more structural realism.
+
+% For instance, both models identify an interesting feature of the traffic patterns in Manhattan, where it is less congested east of Broadway in lower Manhattan than west of Broadway---we speculate that
+
 \subsection{Remaining Puzzles}
+
+% chekcboard?
+% bias?
+% unclear how to decide which one is better
+% improve the softmax one??
+%
 
@@
 
-\begin{itemize}
-\item What happened?
-\item What next?
-\end{itemize}
+%% Feature extraction from parameter values
+
+%TODO
 

### 2017-12-10 reports/paper/paper.tex
.
edited lit rev

@@
 
+Using the New York City Taxi and Limousine Commission (TLC) database of taxi rides, we develop a probabilistic model of transportation in New York City.
+By discretizing the Manhattan geography into a grid structure and by using a mixture model with the latent variable being the route chosen by the driver, we devise a model with a parsimonious number of parameters that is easy to implement even for a large city, fast to estimate, and yields reasonable prediction results relative to benchmarks of Google Maps, linear regression, and neural network.
+Moreover, the parameters of the model can be directly interpreted and visualized as measures of congestion on Manhattan's road network.
 \end{abstract}
@@
 Travel time prediction has historically been a topic of research interest.
-On the topic of freeway travel time prediction, \cite{van2002freeway} proposed a Recurrent Neural Network (RNN) approach in order to address the temporal aspect of traffic prediction; \cite{wu2004travel} adopted the Supportive Vector Regression (SVR) method applied to time-seires analysis and reduced prediction errors in cases where previous methods generate especially large errors.
-We notice that earlier papers share the common feature of using freeway data for training and prediction, which probably arised from the limited availability of GPS data sources.
-Nevertheless, we realize that even for freeways, which in general have less traffic conditions than city roads, there is evidence that there exists a high level of non-linearity in the data, which contributes to the good performance of models such as RNN and SVR.
+On a problem of freeway travel time prediction, \citet*{van2002freeway} proposed a Recurrent Neural Network (RNN) approach in order to address the temporal aspect of traffic prediction; \citet*{wu2004travel} adopted the Supportive Vector Regression (SVR) method applied to time-series analysis and reduced prediction errors in cases where previous methods generate especially large errors.
+We notice that earlier papers share the common feature of using freeway data for training and prediction, which probably arises from the limited availability of GPS data sources.
+Nevertheless, we realize that even for freeways, which in general have less congestion than city roads, there is evidence that there exists a high level of non-linearity in the data, which contributes to the good performance of models such as RNN and SVR.
 
-With the ubiquity of taxi GPS data, recent work focused on estimating the traffic conditions in cities using sparce probe data and used travel time predictions as the metric as the performance of the model.
-\cite{hunter2009path} uses a Bayesian framework and used an expectation maximization algorithm that simultaneously learns the likely paths taken by probe vehicles as well as the travel time distributions through the network; \cite{herring2010estimating} used a Coupled Hidden Markov Model (CHMM) and determines the most likely path by allocating the travel time between two consecutive location observations to roads in the `M' step.
+With the ubiquity of taxi GPS data, recent work focused on estimating the traffic conditions in cities using sparse probe data---a time series of closely spaced GPS location data for each trip---and used travel time predictions as the metric as the performance of the model.
+\citet*{hunter2009path} used a Bayesian framework and used an Expectation-Maximization algorithm that simultaneously learns the likely paths taken by probe vehicles as well as the travel time distributions through the network; \citet*{herring2010estimating} used a Coupled Hidden Markov Model (CHMM) and determines the most likely path by allocating the travel time between two consecutive location observations to roads in the M-step of the EM algorithm.
 
-One thing that distinguishes our attempt to the previous work is that we have much more missing information.
+Compared to previous work in the literature, we have much more missing information.
 Instead of sparse probe data, we are only working with start and end locations, which presents much more challenges in actual path inference and parameter optimization.
-\cite{zhan2013urban} provided some very helpful insights since it used the same TLC dataset of taxi ride.
-It contructed a faithful representation of the manhattan streets network, but limited the discussion to a much smaller district in Manhattan to reduce computational complexity.
+\citet*{zhan2013urban} provided some very helpful insights since it used the same TLC dataset of taxi ride.
+They constructed a faithful representation of the Manhattan streets network, but limited the discussion to a much smaller district in Manhattan to reduce computational complexity.
+Moreover, \citet*{zhan2013urban} also used a softmax regression model for path selection (as detailed in Section~\ref{sec:softmax}), but only restricting the choice set to the 20 shortest paths.
+Our contribution relative to \citet*{zhan2013urban} is twofold: First, our work scales to the entirety of Manhattan by sacrificing the minute realism of their model and discretizing the Manhattan road network as a simply a grid; second, we employs a novel sampling-based approach to estimate the softmax regression model, so as to drastically expand the choice set of drivers in our model.
 
@@
 In principle, it is not difficult to parameterize the prior of $\bm W$ as an undirected graphical model, since we need only to supply edge and unary potentials.
-For instance, to impose a correlated prior on $\bm W$, as suggested by some \cite{hunter2009path}, we simply penalize large differences in neighboring weights in the edge potential, effectively assuming a prior model that is similar to a continuous version of the Ising model.
+For instance, to impose a correlated prior on $\bm W$, as suggested by \citet*{hunter2009path}, we simply penalize large differences in neighboring weights in the edge potential, effectively assuming a prior model that is similar to a continuous version of the Ising model.
 
@@
 \subsection{Softmax Regression Model}
-
+\label{sec:softmax}
 We now consider a more realistic but more complex model, where we parameterize $Z|\bm W$ as a GLM, namely as a softmax regression where $p(Z|\bm W) \propto \exp\pr{-\sum_{i\in Z}w_i}$, encoding drivers' preferences for shorter trips.

### 2017-12-10 reports/paper/paper.tex
intros

@@
 
-% re
+Estimating congestion and predicting trip time are important problems in urban economics, civil engineering, and operations research.
+The problem lends well to approaches in computational probabilistic inference and machine learning, thanks to the increasing availability of GPS data.
+Non-probabilistic machine learning methods, such as neural networks, perform extremely well in prediction, but generate parameters that do not have structural interpretations.
+On the other hand, current probabilistic methods, most of which are based on mixture models, either rely on extremely granular data such as closely spaced GPS measurements \cite{hunter2009path} or are difficult to scale to a large geographical area \cite{zhan2013urban}.
+
+Our main contribution in this project is devising a stylized probabilistic model of travel, congestion, and route choice and creating two strategies of parameterization and approximate inference.
+From a data standpoint, our model does not require extremely granular data---training on a large dataset of taxi rides in New York City that only includes start- and end-trip locations.
+From an implementation standpoint, our model is parsimonious, scalable to large geographical areas and large datasets, and easy to train.
+From a prediction standpoint, our model is close to or outperforms many baseline metrics, such as neural network, linear regression, or Google Maps.
+In stark contrast with black-box methods, our model generates interpretable parameters that directly corresponds to level of congestion in a small area.
+We believe the ability to generate easily interpretable parameters is the foremost of our contributions, as it is transparent and accessible to a broad class of users---leading to, for instance, better urban planning and transportation design by policymakers.
+
+This report proceeds as follows.
+Sections~\ref{sec:backgrd} and \ref{sec:related} briefly discuss some features of the data and some work in the previous literature.
+Section~\ref{sec:model} introduces our modeling framework formally and Section~\ref{sec:inference} introduces two methods for approximate inference of our model.
+Sections~\ref{sec:methods} and \ref{sec:results} discuss training on the TLC dataset and present predictive metrics of the training results relative to baseline.
+Section~\ref{sec:discuss} discusses the results and the accompanying visualization of parameters, while addressing issues of concern and of future inquiry.
+Section~\ref{sec:conc} concludes.
 
 \section{Background}
+\label{sec:backgrd}
 % {
@@
 \section{Related Work}
+\label{sec:related}
 
@@
 \section{Model}
+\label{sec:model}
 
@@
 \section{Methods}
+\label{sec:methods}
 
@@
 \section{Results}
+\label{sec:results}
 
@@
 \section{Discussion}
+\label{sec:discuss}
 
@@
 \section{Conclusion}
+\label{sec:conc}
 

### 2017-12-10 reports/paper/paper.tex
Added puzzles.

@@
 \includegraphics[width=0.5\textwidth]{figs/softmaxmaps.pdf}
-\caption{Learned Weights, Various Times Tuesday. Top Row: Uniform Model.
-Bottom Row: Softmax Model.}
+\caption{Learned weights, various times Tuesday. Top row: uniform model.
+Bottom row: softmax model.}
 \label{fig:unifmap}
@@
 
-% chekcboard?
+While many of our results make sense in the context of the models that produce
+them, some puzzles---both qualitative and quantitative---remain. For instance,
+visual inspection of the grid of weights that the softmax model converges to
+(see bottom row of Figure \ref{fig:unifmap}) reveals a pattern that resembles a
+checkerboard. This behavior is most visible in the 4:00 AM weights, but also
+appears in certain areas of the other grids. We suspect this has to do with the
+superimposition of paths that we train on, but the result largely remains an
+unexpected surprise.
+
+Furthermore, we are also intrigued by the bias our models introduce: both the
+uniform and softmax structural models on average underestimate trip duration
+(see Table \ref{tab:mainres}). Since we optimize an approximate bound, it
+is believable that this procedure introduced bias. Nevertheless, we are
+missing a good intuition as to why we systematically underestimate and whether
+this is a feature or a bug.
+
+\subsection{Future Work}
+
+In thinking of ways to develop our work further, a couple of directions come to
+mind. Foremost among them is taking a closer look at which of the softmax and
+uniform models best captures reality. They perform very similarly on our
+test data, but their weights correspond to structurally different maps of
+Manhattan. Future work could help determine which of these depictions best
+tracks actual traffic conditions.
+
+In addition, our models currently train from scratch on each partition of the
+data. However, it is not hard to imagine that, for example, traffic conditions
+at 8:00 AM on Tuesdays can tell you something about conditions at 9:00 AM on
+Tuesdays, or 8:00 AM on Wednesdays. Exploring ways to leverage this structure
+through transfer learning could significantly speed up the training of our
+models.
+
+% checkboard?
 % bias?
@@
 % improve the softmax one??
+% transfer learning from different times
+
+% \begin{itemize}
+% \item What conclusions can you draw from the results section?
+% \item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
 % 
-
-\begin{itemize}
-\item What conclusions can you draw from the results section?
-\item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
-
-\end{itemize}
+% \end{itemize}
 
@@
 %% Feature extraction from parameter values
-
-%TODO
 

### 2017-12-10 reports/paper/paper.tex
tweaks

@@
 constraint on how many test data points it could be tested on. We assess its
-performance on 2,500 data points in a representative partioned test dataset.
+performance on approximately 2,500 data points in a representative partioned test dataset.
+Moreover, Google Maps API does not allow predicting travel time in the past, and thus for each trip, we use a random Tuesday 8:00 AM in January 2018 as a proxy.
+From our tests, the predictions of Google Maps API are not significantly different, whether the user submits a start time or not.
+Thus, we believe that this is a reasonable approximation given limitations to the data.
 
@@
 The softmax regression model was trained
-using the Adam optimizer \cite{kingma2014adam} with learning rate $10^{-2}$ for 10 epochs,
-taking approximately an hour.
+using the Adam optimizer \cite{kingma2014adam} with learning rate $5\times 10^{-3}$ for various epochs for different data-subsets,\footnote{The epochs are calculated so that the total numbers of trips trained by the model are approximately constant across the subsets of data.
+This means that we trained for 131 epochs for 4:00 AM but only 6 epochs for 8:00 AM, since the two data-subsets have drastically different number of trips.
+Nonetheless, as Figure~\ref{fig:softmaxloss} shows, we reached convergence in each dataset we trained on.}
+taking approximately thirty minutes for each data-subset.
 Additionally, the softmax regression model take hyperparameters $L$ (sample size in each stochastic draw) and $\sigma^2$ (variance of trip time).

### 2017-12-10 reports/paper/paper.tex
Add conclusion.

@@
 
-\subsection{Future Work}
-
-In thinking of ways to develop our work further, a couple of directions come to
-mind. Foremost among them is taking a closer look at which of the softmax and
-uniform models best captures reality. They perform very similarly on our
-test data, but their weights correspond to structurally different maps of
-Manhattan. Future work could help determine which of these depictions best
-tracks actual traffic conditions.
-
-In addition, our models currently train from scratch on each partition of the
-data. However, it is not hard to imagine that, for example, traffic conditions
-at 8:00 AM on Tuesdays can tell you something about conditions at 9:00 AM on
-Tuesdays, or 8:00 AM on Wednesdays. Exploring ways to leverage this structure
-through transfer learning could significantly speed up the training of our
-models.
-
 % checkboard?
@@
 
+In this paper, we introduce a model that scales effectively, can be trained
+without granular data, performs competitively with baselines (including Google
+Maps predictions), and has parameters that correspond to structural measures of
+congestion. We stress the combination of these attributes as the value of our
+model, since various other models exist if we relax these criteria. For
+instance, \citealt{hunter2009path} trains on granular data, 
+\citealt{zhan2013urban} scales poorly, and a neural network does not have
+structural parameters.
+
+\subsection{Future Work}
+
+In thinking of ways to develop our work further, a couple of directions come to
+mind. Foremost among them is taking a closer look at which of the softmax and
+uniform models best captures reality. They perform very similarly on our
+test data, but their weights correspond to structurally different maps of
+Manhattan. Future work could help determine which of these depictions best
+tracks actual traffic conditions.
+
+In addition, our models currently train from scratch on each partition of the
+data. However, it is not hard to imagine that, for example, traffic conditions
+at 8:00 AM on Tuesdays can tell you something about conditions at 9:00 AM on
+Tuesdays, or 8:00 AM on Wednesdays. Exploring ways to leverage this structure
+through transfer learning could significantly speed up the training of our
+models.
+
 %% Feature extraction from parameter values

### 2017-12-10 reports/paper/paper.tex
edits

@@
 Using the New York City Taxi and Limousine Commission (TLC) database of taxi rides, we develop a probabilistic model of transportation in New York City.
-By discretizing the Manhattan geography into a grid structure and by using a mixture model with the latent variable being the route chosen by the driver, we devise a model with a parsimonious number of parameters that is easy to implement even for a large city, fast to estimate, and yields reasonable prediction results relative to benchmarks of Google Maps, linear regression, and neural network.
+By discretizing the Manhattan geography into a grid structure and by using a mixture model with the latent variable being the route chosen by the driver, we devise a model with a parsimonious number of parameters.
+Our model is easy to implement even for a large city, fast to estimate, and yields reasonable prediction results relative to benchmarks of Google Maps, linear regression, and neural network.
 Moreover, the parameters of the model can be directly interpreted and visualized as measures of congestion on Manhattan's road network.
@@
 From an implementation standpoint, our model is parsimonious, scalable to large geographical areas and large datasets, and easy to train.
-From a prediction standpoint, our model is close to or outperforms many baseline metrics, such as neural network, linear regression, or Google Maps.
+From a prediction standpoint, our model is close to or outperforms many baseline metrics, such as neural networks, linear regression, or Google Maps.
 In stark contrast with black-box methods, our model generates interpretable parameters that directly corresponds to level of congestion in a small area.
@@
 characterized by the time and location (latitude and longitude) of the pickup as
-well as the corresponding quantities for the drop-off\footnote{Additional
+well as the corresponding quantities for the drop-off.\footnote{Additional
 information orthogonal to predicting trip duration---such as passenger
-counts and tip---is discarded}. Moreover, there is an abundance of data. In
+counts and tip---is discarded} Moreover, there is an abundance of data. In
 January of 2009 alone, there were over 14 million taxi rides recorded.
@@
 reasonable for a taxi trip, just over 1\% of trips have a drop-off time that is
-before the pick-up time, making for a trip with negative duration. Furthermore,
+\emph{before} the pick-up time, making for a trip with negative duration. Furthermore,
 the longest trip has duration of just over 42 days. It is difficult to
@@
 Additionally, whereas we would like to infer road conditions along the route of
-a taxi trip, we can only observe the start and end-points. The fact that the
-route taken is unobservable is primarily responsible for our inference
-challenges.
+a taxi trip, we can only observe the start- and end-points. Therefore, our main inference challenge is the latent route choice of each trip.
 
@@
 Travel time prediction has historically been a topic of research interest.
-On a problem of freeway travel time prediction, \citet*{van2002freeway} proposed a Recurrent Neural Network (RNN) approach in order to address the temporal aspect of traffic prediction; \citet*{wu2004travel} adopted the Supportive Vector Regression (SVR) method applied to time-series analysis and reduced prediction errors in cases where previous methods generate especially large errors.
-We notice that earlier papers share the common feature of using freeway data for training and prediction, which probably arises from the limited availability of GPS data sources.
-Nevertheless, we realize that even for freeways, which in general have less congestion than city roads, there is evidence that there exists a high level of non-linearity in the data, which contributes to the good performance of models such as RNN and SVR.
+On a problem of freeway travel time prediction, \citet*{van2002freeway} proposed a Recurrent Neural Network (RNN) approach in order to address the temporal aspect of traffic prediction; \citet*{wu2004travel} adopted the Support Vector Regression (SVR) method applied to time-series analysis and reduced prediction errors in cases where previous methods generate especially large errors.
+We notice that earlier papers used freeway data for training and prediction, which probably arises from the limited availability of GPS data sources.
+Nevertheless, we realize that even for freeways, which in general have less congestion than city roads, there is evidence of non-linearity in the data, which contributes to the good performance of models such as RNN and SVR.
 
-With the ubiquity of taxi GPS data, recent work focused on estimating the traffic conditions in cities using sparse probe data---a time series of closely spaced GPS location data for each trip---and used travel time predictions as the metric as the performance of the model.
-\citet*{hunter2009path} used a Bayesian framework and used an Expectation-Maximization algorithm that simultaneously learns the likely paths taken by probe vehicles as well as the travel time distributions through the network; \citet*{herring2010estimating} used a Coupled Hidden Markov Model (CHMM) and determines the most likely path by allocating the travel time between two consecutive location observations to roads in the M-step of the EM algorithm.
+With the ubiquity of GPS data, recent work focuses on estimating the traffic conditions in cities using sparse probe data---a time series of closely spaced GPS location data for each trip---and uses travel time predictions as the metric as the performance of the model.
+\citet*{hunter2009path} used a Bayesian framework and used an Expectation-Maximization algorithm that simultaneously learns the likely paths taken by probe vehicles as well as the travel time distributions through the network; \citet*{herring2010estimating} used a Coupled Hidden Markov Model (CHMM) and determined the most likely path by allocating the travel time between two consecutive location observations to roads in the M-step of the EM algorithm.
 
 Compared to previous work in the literature, we have much more missing information.
-Instead of sparse probe data, we are only working with start and end locations, which presents much more challenges in actual path inference and parameter optimization.
-\citet*{zhan2013urban} provided some very helpful insights since it used the same TLC dataset of taxi ride.
+Instead of sparse probe data, we are only working with start and end locations, which presents challenges in actual path inference and parameter optimization.
+\citet*{zhan2013urban} provide some helpful insights since they used the same TLC dataset of taxi ride.
 They constructed a faithful representation of the Manhattan streets network, but limited the discussion to a much smaller district in Manhattan to reduce computational complexity.
 Moreover, \citet*{zhan2013urban} also used a softmax regression model for path selection (as detailed in Section~\ref{sec:softmax}), but only restricting the choice set to the 20 shortest paths.
-Our contribution relative to \citet*{zhan2013urban} is twofold: First, our work scales to the entirety of Manhattan by sacrificing the minute realism of their model and discretizing the Manhattan road network as a simply a grid; second, we employs a novel sampling-based approach to estimate the softmax regression model, so as to drastically expand the choice set of drivers in our model.
+Our contribution relative to \citet*{zhan2013urban} is twofold: First, our work scales to the entirety of Manhattan by sacrificing the minute realism of their model and discretizing the Manhattan road network as a simply a grid; second, we employ a novel sampling-based approach to estimate the softmax regression model, so as to drastically expand the choice set of drivers in our model.
 
@@
 \]
-in order to encode the fact that drivers avoid routes that take a long period of time.
-In the \emph{uniform} model, we simply assume that route choice is independent and uniform on the set of reasonable paths: \[
+in order to encode the fact that drivers avoid routes that take a long period of time.\footnote{In general, we can have a \emph{temperature} parameter to the softmax that is not $1$.
+However, this is a hyperparameter since route choice is latent and thus not observable.
+To simplify notation and to be consistent with the applied parts of our project, we use softmax with temperature $1$.} In the \emph{uniform} model, we simply assume that route choice is independent and uniform on the set of reasonable paths: \[
 p(Z\uppr{k}|\bm W) \propto 1.
@@
 \end{align*}
-The expectation is a sum of the size $|\bm Z\uppr{k}|,$ which, for an $m\times n$ trip\footnote{By an $m\times n$ trip, we mean a trip with east-west distance $n$ and north-south distance $m$}, is of $\binom{m+n}{n} \approx \frac{(n+m)^n}{n^n}e^n$ terms.
-Computing this expectation is the main inference challenge of our project.
+The expectation is a sum of the size $|\bm Z\uppr{k}|,$ which, for an $m\times n$ trip\footnote{By an $m\times n$ trip, we mean a trip with east-west distance $n$ and north-south distance $m$}, is of $\binom{m+n}{n} \approx \frac{(n+m)^n}{n^n}e^n$ terms.\footnote{Sterling's approximation} Computing this expectation is the main inference challenge of our project.
 
@@
 \]
-and computing the expectation requires the conditional distribution $p(Z | T, \bm W) = \frac{p(T|Z,\bm W)p(Z|\bm W)}{p(T|\bm W)}$, where the denominator is intractable to compute.
+and computing the expectation requires the conditional distribution \[p(Z | T, \bm W) = \frac{p(T|Z,\bm W)p(Z|\bm W)}{p(T|\bm W)},\] where the denominator is intractable to compute.
 We might also consider techniques in variational inference \cite{blei2017variational}.
 However, the latent variable $Z\uppr{k}$ is a random set of indices with the property that the indices form a path on the grid, and thus cannot be reasonably decomposed into independent components, ruling out the mean-field algorithm.
-Another promising option is stochastic gradient variational Bayes (SGVB) and variational autoencoders \cite{kingma2013auto}, which is designed for large datasets for which the EM algorithm fails.
+Another promising option is stochastic gradient variational Bayes (SGVB) and variational autoencoders \cite{kingma2013auto}, which are designed for large datasets on which the EM algorithm fails.
 However, SGVB requires a reparameterization of the latent variable drawn from an approximate distribution, in order for the gradients to be properly computed.
@@
 This is the same technique as in importance sampling, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations.
-Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields the following upper bound for negative log-likelihood, up to a constant,
+We note that the same trick allows us to replace any sum over $\bm Z$ with a weighted sum over the uniformly chosen sample.
+Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields a lower bound for log-likelihood, up to a constant \[
+-\frac{1}{L}\bk{\sum_{\tilde Z} \frac{1}{2\sigma^2} \pr{T - \sum_{i\in \tilde Z} w_i}^2 + \log \frac{\exp\pr{-\sum_i w_i}}{\sum_{\bm Z}\exp\pr{-\sum_i w_i}}}.
+\]
+Using $\tilde Z$ again to approximate the sum in the denominator yields the following upper bound for negative log-likelihood, up to a constant,
 \begin{align}
@@
 Computing prediction is slightly more subtle in the softmax regression model.
-In the softmax regression model, we sample $L_{\text{pred}}$ uniformly chosen paths and take the weighted average of the sums of $w_i$ in each path, where the weights are proportional to $\exp(-\sum_{i \in Z} w_i)$ for path $Z$.
+In the softmax regression model, we sample $L_{\text{pred}}$\footnote{In the implementation and application, we used $L_\text{pred} = 50$.} uniformly chosen paths and take the weighted average of the sums of $w_i$ in each path, where the weights are proportional to $\exp(-\sum_{i \in Z} w_i)$ for path $Z$.
 

### 2017-12-10 reports/paper/paper.tex
Paper edits second half.

@@
 
-In addition to adding columns, we filtered the original dataset. We only consider trips
-whose start- and end-locations that falls inside a rectangle that bounds Manhattan. We also removed trips with duration less than two minutes or more than two hours.
+In addition to adding columns, we filtered the original dataset. We only
+considered trips
+whose start- and end-locations fall inside a rectangle that bounds Manhattan. We also removed trips with duration less than two minutes or more than two hours.
 
 We performed a principal component analysis on all de-meaned and
-standardized geographic coordinates (both start
-and end locations) from the filtered data.
-We thus transform the coordinates into the first and second principal components.
-This allowed us to compute the rotational matrix so that Manhattan lies approximately orthogonal to the $x$-axis.
-For the discrete
+standardized geographic coordinates (both start-
+and end-locations) from the filtered data. We transformed the coordinates into
+the first and second principal components. This allowed us to compute the
+rotational matrix so that the major axis of Manhattan lies approximately orthogonal to the $x$-axis. For the discrete
 grid models, these coordinates were further discretized into a $20 \times 70$
@@
 trained on the entire dataset (e.g.\ Neural Network), and some were trained
-individually on the partioned datasets (e.g.\ Structural Models).
+individually on the partioned datasets (e.g.\ our structural models).
 
@@
 We calculated three baselines. The first was a linear regression predicting
-trip duration from trip distance (calculated using the Pythagorean Theorem and
-the start and end coordinates), and trained on a partitioned data set.
+trip duration from Euclidian distance, and trained on a partitioned data set.
 
@@
 While there was no need to train this baseline, there was a rate limit
-constraint on how many test data points it could be tested on. We assess its
-performance on approximately 2,500 data points in a representative partioned test dataset.
-Moreover, Google Maps API does not allow predicting travel time in the past, and thus for each trip, we use a random Tuesday 8:00 AM in January 2018 as a proxy.
+constraint on how many test data points it could be tested on. We assessed its
+performance on approximately 2,500 data points in a representative partioned
+test dataset. Moreover, Google Maps API does not allow predicting travel time in
+the past, and thus for each trip, we used a random Tuesday 8:00 AM in January 2018 as a proxy.
 From our tests, the predictions of Google Maps API are not significantly different, whether the user submits a start time or not.
@@
 Training implementations varied from model to model. In training the neural
-network baseline, SGD was used with a learning rate of $10^{-7}$ run for 12
+network baseline, we used SGD with a learning rate of $10^{-7}$ run for twelve 
 epochs on the full data set. This took between 30 minutes and an hour. The
@@
 dataset corresponding to all trips on a given day of the week and on a given
-hour. Convergence appeared after roughly two epochs, and the model was run for 5
-epochs taking just under ten minutes.
+hour. Convergence appeared after roughly two epochs, and the model was run for
+five epochs taking just under ten minutes.
 
@@
 Nonetheless, as Figure~\ref{fig:softmaxloss} shows, we reached convergence in each dataset we trained on.}
-taking approximately thirty minutes for each data-subset.
-Additionally, the softmax regression model take hyperparameters $L$ (sample size in each stochastic draw) and $\sigma^2$ (variance of trip time).
-We chose $L = 20$ and $\sigma^2 = 100$.
-These values are again chosen to strike a bargain between realism and computation feasibility.
-The model appears to converge by the second epoch.
+taking approximately thirty minutes for each data-subset. Additionally, the
+softmax regression model takes hyperparameters $L$ (sample size in each
+stochastic draw) and $\sigma^2$ (variance of trip time). We chose $L = 20$ and
+$\sigma^2 = 100$. These were again chosen to strike a bargain between realism
+and computation feasibility. The model appeared to converge by the second epoch.
 
@@
 
-We train our model and test the model on an out-of-sample test-set using the methods discussed in Section~\ref{sub:predict}.
-As a simple check of the training process, we plot in Figure~\ref{fig:softmaxloss} the loss obtained when fitting the softmax regression model.
-We see that the model converges relatively fast.\footnote{The full fitting takes approximately 30 minutes per dataset (2.9 GHz Intel Core i5, 8GB RAM), but the model appears to converge in only 20\% of that time.} The convergence plot for the uniform model is extremely similar and thus omitted.
+We train our model and test the model on an out-of-sample test-set using the
+methods discussed in Section~\ref{sub:predict}. As a simple check of the
+training process, we plot in Figure~\ref{fig:softmaxloss} the loss obtained when
+fitting the softmax regression model. We see that the model converges relatively
+fast.\footnote{The full fitting takes approximately 30 minutes per dataset (2.9
+GHz Intel Core i5, 8GB RAM), but the model appears to converge in only one-fifth
+of this time.} The convergence plot for the uniform model is extremely similar and thus omitted.
 Table~\ref{tab:mainres} contains the main prediction results of our model relative to some benchmark predictions
@@
 We note that, in terms of predictive metrics, our models were able to outperform Google Maps and linear regression, but were not able to outperform the neural network, even though the difference is not very large.
-Despite the disappointing outcome, the advantage of our framework is that the parameters have structural interpretations, whereas the neural network is a black box.
 
@@
 
-Table~\ref{tab:mainres} is encouraging.
-Our models successfully leverage the power of geography and picks up more signal in the data than distance of the trip alone, as demonstrated by the significantly improved performance over linear regression.
-Perhaps more strikingly, our model slightly outperforms the predictions generated by Google Maps, which means that the predictive accuracy of the model is on par with the standard of a widely used consumer product.
-Even though we were not able to outperform the neural network, we reiterate that the main advantage of our approach is to have interpretable parameters, which are visualized in Figure~\ref{fig:unifmap}.
-We turn to the figure in the next section.
+Table~\ref{tab:mainres} is encouraging. Our models successfully leverage the
+power of geography and pick up more signal in the data than distance of the trip
+alone, as demonstrated by the significantly improved performance over linear
+regression. Perhaps more strikingly, our model slightly outperforms the
+predictions generated by Google Maps, which means that the predictive accuracy
+of the model is on par with the standard of a widely used consumer product. Even
+though we were not able to outperform the neural network, we reiterate that the
+main advantage of our approach is to have interpretable parameters, which are
+visualized in Figure~\ref{fig:unifmap}. We turn to the figure now.
 % \begin{itemize}
@@
 \caption{Learned weights, various times Tuesday. Top row: uniform model.
-Bottom row: softmax model.}
+Bottom row: softmax model. \emph{In the grids, the southern border of Central Park lies
+approximately three-sevenths of the way up}}
 \label{fig:unifmap}
@@
 We identify that 4:00 AM is indeed less congested than rush hour, that Midtown Manhattan seems to be particularly congested during the day, and that the congested areas have an average speed of 4--6 miles per hour.
-The visualization also confirms some less-than-well-known features; the areas that demonstrate congestion at 4:00 AM largely confirms with anecdotal evidence.\footnote{According to \textit{Business Insider} (\url{http://read.bi/164vSK1}), East Village and surrounding areas rank highly for nightlife.
+The visualization also confirms some less-than-well-known features; the areas that demonstrate congestion at 4:00 AM largely confirms anecdotal evidence.\footnote{According to \textit{Business Insider} (\url{http://read.bi/164vSK1}), East Village and surrounding areas rank highly for nightlife.
 %#GETLIT
@@
 appears in certain areas of the other grids. We suspect this has to do with the
-superimposition of paths that we train on, but the result largely remains an
-unexpected surprise.
+superimposition of paths that we train on, but the result largely remains a
+surprise.
 
@@
 missing a good intuition as to why we systematically underestimate and whether
-this is a feature or a bug.
+this is a feature or a bug. It is also interesting to note that the Google Maps
+predictions exhibit a bias in the opposite direction.
 
@@
 
-In thinking of ways to develop our work further, a couple of directions come to
-mind. Foremost among them is taking a closer look at which of the softmax and
+We now discuss a few directions for future work. Foremost among them is
+taking a closer look at which of the softmax and
 uniform models best captures reality. They perform very similarly on our
@@
 models.
+
+Finally, we note that our models condition only on hour and day of the week when
+making predictions. However, our framework can incorporate a wider range of
+relevant information. For instance, if weather data were accessible, the effect
+of different weather conditions on road congestion could be measured. Similarly,
+abnormal traffic conditions on holidays could be identified.
 

### 2017-12-10 reports/paper/paper.tex
QED

@@
 where $\tilde Z$ is drawn uniformly from $\bm Z$.
-This is the same technique as in importance sampling, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations.
+This is the same technique as in importance sampling \cite{gelman2014bayesian}, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations.
 We note that the same trick allows us to replace any sum over $\bm Z$ with a weighted sum over the uniformly chosen sample.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{figs/unifmaps.pdf}
+\includegraphics[width=0.5\textwidth]{figs/softmaxmaps.pdf}
+\caption{Learned weights, various times Tuesday. Top row: uniform model.
+Bottom row: softmax model.}
+\label{fig:unifmap}
+\end{figure}
+
 \section{Discussion}
 \label{sec:discuss}
-
-% table
-%
 
@@
 main advantage of our approach is to have interpretable parameters, which are
-visualized in Figure~\ref{fig:unifmap}. We turn to the figure now.
+visualized in Figure~\ref{fig:unifmap}.\footnote{In Figure~\ref{fig:unifmap}, the southern border of Central Park lies
+approximately three-sevenths of the way up} We turn to the figure now.
 % \begin{itemize}
@@
 % \end{itemize}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.5\textwidth]{figs/unifmaps.pdf}
-\includegraphics[width=0.5\textwidth]{figs/softmaxmaps.pdf}
-\caption{Learned weights, various times Tuesday. Top row: uniform model.
-Bottom row: softmax model. \emph{In the grids, the southern border of Central Park lies
-approximately three-sevenths of the way up}}
-\label{fig:unifmap}
-\end{figure}
 
@@
 relevant information. For instance, if weather data were accessible, the effect
-of different weather conditions on road congestion could be measured. Similarly,
+of different weather conditions on road congestion could be measured by adding the weather covariates to the Normal mean. Similarly,
 abnormal traffic conditions on holidays could be identified.
